---
title: "PDAC-A stratification"
author: "Marc Elosua-Bayes"
date: "5/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this Rmarkdown document we will divide the spots into cancer and non-cancer and then check if there is region-specific immune cell enrichment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(NMF)
library(purrr)
library(tibble)
library(ggplot2)
library(Matrix)
library(Seurat)
library(Spaniel)
library(SPOTlight)
library(flextable)
source("utils/bin.r")
source("utils/spatial_plot_spaniel.R")
```
Paths
```{r}
#### Setting vrs ####
tech <- "sc"
tissue <- "pdac_unpaired"
dwn_smplng <- "both"
org <- "hs"
source("misc/paths_vrs.R")

clust_vr <- "cell_type_specific"
cl_n <- 100
method <- "nsNMF"
transf <- "uv"
hvg <- 3000
ntop <- NULL
FC <- "all"
pct1 <- "all"

id_nmf <- sprintf("cln-%s_transf-%s_method-%s_hvg-%s_FC-%s_pct1-%s", 
                  cl_n, transf, method, hvg, FC, pct1)


library(RColorBrewer)
n <- 60
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector  <-  unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))


# sample dictionary
sample_dict <- list()
sample_dict[["GSM3036911"]] = "PDAC-A ST1"
sample_dict[["GSM3405534"]] = "PDAC-B ST1"
sample_dict[["GSM4100721"]] = "PDAC-A ST2"
sample_dict[["GSM4100722"]] = "PDAC-A ST3"
sample_dict[["GSM4100723"]] = "PDAC-B ST2"
sample_dict[["GSM4100724"]] = "PDAC-B ST3"

pdac_a_smpl <- c("GSM3036911", "GSM4100721", "GSM4100722")
pdac_b_smpl <- c("GSM3405534", "GSM4100723", "GSM4100724")
pdac_smpl <- c(pdac_a_smpl, pdac_b_smpl)

dir.create(path = sprintf("%s/%s", an_pdac, plt_dir), showWarnings = FALSE, recursive = TRUE)
dir.create(path = sprintf("%s/%s", an_pdac, robj_dir), showWarnings = FALSE, recursive = TRUE)
```


## Load data
### scRNAseq data
```{r}
# ica_se <- readRDS("data/immune_cancer_atlas/pancreas_new_celltypes_100.rds")
# ica_se <- ica_se[, !ica_se$specific_cell_type %in% c("Macro_2", "Macro_3")]

update_annot <- readr::read_csv(
  "data/immune_cancer_atlas/TICAtlas_metadata.csv"
  ) %>% dplyr::rename(barcode = X1)
```

### Spatial data
```{r}
st_list <- sapply(pdac_smpl, USE.NAMES = TRUE, function(geo) {
  print(geo)
  sample <- stringr::str_split(string = sample_dict[[geo]], pattern = " ")[[1]][1]
  geo_path <- sprintf("analysis/pancreas_PDAC/data/%s/%s", sample, geo)
  st_tibble <- readr::read_tsv(file = sprintf("%s/filtered_feature_bc_matrix.tsv", geo_path), 
                            col_names = TRUE) %>%
    tibble::column_to_rownames("X1")
  
  st_se <- Spaniel::createSeurat(counts = st_tibble, 
               barcodeFile = sprintf("%s/spatial/tissue_positions_list_spatial_object.tsv", geo_path), 
               projectName = "PDAC-B",
               sectionNumber = "1")
  
  ### Load histological image into R
  fn_image <- list.files(path = sprintf("%s/spatial", geo_path), pattern = "lowres_rot.jpg$")[1]
  imgFile <- sprintf("%s/spatial/%s", geo_path, fn_image)
  image <- Spaniel::parseImage(imgFile)
  st_se@images <- list(image)
  
  st_se$slice <- sample_dict[[geo]]
  return(list(st_se))
})
```

### Deconvolution matrices
Load the Paired deconvolution matrices so that we can get the labels of tumoral vs non-tumoral spots
```{r}
decon_mtrx_dir <- sprintf("%s/%s", an_pdac, robj_dir)
decon_fn_ls <- list.files(path = decon_mtrx_dir,
                       pattern = "^decon_mtrx_paired")

decon_fn_ls <- decon_fn_ls[grepl(pattern = "itai_2",
                                 x = decon_fn_ls)]


paired_decon_mtrx_ls <- lapply(decon_fn_ls, function(fn) {
  path_str <- paste(decon_mtrx_dir, fn, sep = "/")
  decon_mtrx <- readRDS(file = path_str)
  return(decon_mtrx)
})

names_ls <- sapply(decon_fn_ls, USE.NAMES = FALSE, function(i) stringr::str_split(string = i, pattern = "_")[[1]][4])

names(paired_decon_mtrx_ls) <- names_ls
```

Load the immune cell atlas matrices to have the immune cells mapped onto the tissue
```{r}
decon_mtrx_dir <- sprintf("%s/%s", an_pdac, robj_dir)
decon_fn_ls <- list.files(path = decon_mtrx_dir,
                       pattern = "^decon_mtrx_ica-integration_specific")[1:6]



immune_decon_mtrx_ls <- lapply(decon_fn_ls, function(fn) {
  path_str <- paste(decon_mtrx_dir, fn, sep = "/")
  decon_mtrx <- readRDS(file = path_str)
  return(decon_mtrx)
})

names_ls <- sapply(decon_fn_ls, USE.NAMES = FALSE, function(i) stringr::str_split(string = i, pattern = "_")[[1]][5])

names(immune_decon_mtrx_ls) <- names_ls

```

### Color dataframe
Cell type specific color
```{r}
cell_types_plt <- sort(unique(update_annot$cell_type))

pdac_plt_names <- data.frame(df_name = gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
                                            x = cell_types_plt,
                                            perl = TRUE), 
                           plt_name = cell_types_plt,
                           col_ct = col_vector[seq_len(length(cell_types_plt))])

```

Stratification specific color
```{r}
strat_plt_col <- data.frame(strat = c("Intermediate", "Non-Tumoral", "Removed manually", "Tumoral"),
                            strat_col = c("#4DBBD5FF", "#E64B35FF", "lightgrey", "#00A087FF"),
                            stringsAsFactors = FALSE)
```

### Visualization
```{r}
st_se <- st_list[[1]]
```

### Classify by cancer-cell proportion
Look at the best threshold to divide cells by tumor/no tumor
```{r}
decon_mtrx_paired <- paired_decon_mtrx_ls[[1]]

head(decon_mtrx_paired)
cancer_prop <- rowSums(decon_mtrx_paired[, c("Cancer.clone.(S100A4)", "Cancer.clone.(TM4SF1)")])
healthy_prop <- rowSums(decon_mtrx_paired[, c("Acinar.cells", "Centroacinar.ductal.cells", "Terminal.ductal.cells", "Antigen-presenting.ductal.cells", "Tuft.cells" )])

# par(mfrow = c(2, 1))
st_se[["cancer_prop"]] <- cancer_prop
st_se[["healthy_prop"]] <- healthy_prop
```

```{r}
pt1 <- Spaniel::spanielPlot(object = st_se,
                     grob = st_se@images[[1]], 
                     plotType = "Cluster", 
                     clusterRes = "cancer_prop", 
                     ptSize = 5) +
  theme_void() +
  coord_fixed(1) +
  scale_alpha(range = c(1, 1)) +
  scale_color_gradientn(
        colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"))
        # colours = heat.colors(10, rev = TRUE))

pt2 <- Spaniel::spanielPlot(object = st_se,
                     grob = st_se@images[[1]], 
                     plotType = "Cluster", 
                     clusterRes = "healthy_prop", 
                     ptSize = 5) +
  theme_void() +
  coord_fixed(1) +
  scale_alpha(range = c(1, 1)) +
  scale_color_gradientn(
        colours = wesanderson::wes_palette("Zissou1", 100, type = "continuous"))
        # colours = heat.colors(10, rev = TRUE))

pt3 <- ggplot() +
  geom_histogram(aes(x = cancer_prop), binwidth = 0.01) +
  geom_vline(xintercept = 0.25, col = "red")

ggpubr::ggarrange(pt1, pt2, pt3, ncol = 3)

```

### Tissue stratification
According to the density plot and the visualization we can set an accurate threshold to differentiate tumoral and non-tumoral cells at 0.4.
```{r}
st_se[["status_2_territories"]] <- if_else(st_se$cancer_prop > 0.4, "Tumoral", "Non-Tumoral")

# st_se[["status_3_territories"]] <- if_else(st_se$cancer_prop < 0.4 & 
#                                        st_se$healthy_prop < 0.4 & 
#                                        st_se$y > 16, "Intermediate",
#                                      if_else(st_se$cancer_prop > 0.4, "Tumoral", "Non-Tumoral"))

st_se[["status_3_territories"]] <- if_else(st_se$cancer_prop < 0.4 &
                                             st_se$healthy_prop < 0.4 & 
                                             st_se$y > 16,
                                           "Intermediate",
                                           if_else(st_se$cancer_prop < 0.4 &
                                             st_se$healthy_prop < 0.4 & 
                                             st_se$y < 16,
                                           "Removed manually",
                                           if_else(st_se$cancer_prop > 0.4,
                                                   "Tumoral", 
                                                   "Non-Tumoral")))

```


Visualize on the slide health_status_2_territories
```{r}
# Inverse Y to flip the coordinates
data_df <- data.frame(st_se@meta.data)
data_df$y_inv <- 36 - data_df$y

# data_df[, point_size] <- if_else(data_df[, point_size] == 0, NA_real_, data_df[, point_size])

tmp_plt_2 <- ggplot(data_df,
                ggplot2::aes_string("x", "y_inv",
                                    color = "status_2_territories" 
                                    # alpha = point_alpha
                                    )) +
  ggplot2::xlim(1, 33) +
  ggplot2::ylim(1, 35) +
  # Layer 1 - Plot image
  ggplot2::annotation_custom(st_se@images[[1]],
                             xmin = 1,
                             xmax = 33,
                             ymin = 1,
                             ymax = 35) +
  # Layer 2 - Plot points
  geom_point(size = 5, alpha = 0.8) +
  labs(color = "Tissue stratification") +
  coord_fixed(1) +
  theme_void() +
  scale_color_manual(values = c("#00A087FF", "#E64B35FF"))

tmp_plt_2
```

Visualize on the slide health_status_3_territories
```{r}
# Inverse Y to flip the coordinates
point_size <- "status_3_territories"
point_colour <- "status_3_territories"
data_df <- data.frame(st_se@meta.data)
data_df$y_inv <- 36 - data_df$y

# data_df[, point_size] <- if_else(data_df[, point_size] == 0, NA_real_, data_df[, point_size])

ind_3 <- which(strat_plt_col$strat == sort(unique(data_df$status_3_territories)))
plt3_col <- sort(strat_plt_col[ind_3, "strat_col"])
  
tmp_plt_3 <- ggplot2::ggplot(data_df,
                ggplot2::aes_string("x", "y_inv",
                                    color = "status_3_territories" 
                                    # alpha = point_alpha
                                    )) +
  ggplot2::xlim(1, 33) +
  ggplot2::ylim(1, 35) +
  # Layer 1 - Plot image
  ggplot2::annotation_custom(st_se@images[[1]],
                             xmin = 1,
                             xmax = 33,
                             ymin = 1,
                             ymax = 35) +
  # Layer 2 - Plot points
  geom_point(size = 5) +
  coord_fixed(1) +
  theme_void() +
  scale_color_manual(values = strat_plt_col[strat_plt_col$strat %in% unique(data_df$status_3_territories), "strat_col"])
tmp_plt_3
```

Get capture location identifiers to assign tumoral/non-tumoral
```{r}
# tumor_id <- colnames(st_se)[, st_se$health_status == 1]
```

#### Joining immune cell type information
Transfer assignment to Immune cell mapping
```{r}
decon_mtrx_immune <- immune_decon_mtrx_ls[[1]][[2]]

decon_mtrx_immune[decon_mtrx_immune > 0] <- 1

decon_df_immune <- data.frame(decon_mtrx_immune)

decon_df_immune$status_2_territories <- st_se@meta.data[, "status_2_territories"]
```

#### Analyzing status_2_territories

Proportion heatmap
```{r}
tmp_df <- decon_df_immune %>%
  dplyr::select(-c("res_ss")) %>% 
  dplyr::mutate(total = 1) %>% 
  dplyr::group_by(status_2_territories) %>% 
  dplyr::summarise_if(is.numeric, ~sum(.)) %>% 
  data.frame() %>% 
  tibble::column_to_rownames("status_2_territories")

prop_df <- tmp_df[, 1:(ncol(tmp_df) - 1)] / tmp_df[, ncol(tmp_df)]
tmp_dif <- sapply(colnames(prop_df), function(i) prop_df[1, i] - prop_df[2, i])
```

Get difference function
```{r}
group_prop_differences <- function(df, grp_vr) {
  tmp_df <- df %>%
  # dplyr::select(-c("status_3_territories", "res_ss")) %>% 
  dplyr::mutate(total = 1) %>% 
  dplyr::group_by_at(grp_vr) %>% 
  dplyr::summarise_if(is.numeric, ~sum(.)) %>% 
  data.frame() %>% 
  tibble::column_to_rownames(grp_vr)

  prop_df <- tmp_df[, 1:(ncol(tmp_df) - 1)] / tmp_df[, ncol(tmp_df)]
  tmp_dif <- sapply(colnames(prop_df), function(i) prop_df[1, i] - prop_df[2, i])
  
  return(tmp_dif)
}

```

Permutation test
```{r}
set.seed(123)

decon_df_immune2 <- decon_df_immune %>%
  dplyr::select(-"res_ss")

# Get difference distributions
perm_diff <- sapply(1:10000, function(iter){
  
  shuff_df <- as.data.frame(lapply(decon_df_immune2, sample)) %>% 
    mutate(status_2_territories = decon_df_immune2$status_2_territories)
  
  tmp <- group_prop_differences(df = shuff_df, grp_vr = "status_2_territories")
  return(tmp)
})

# Get double tailed pvalue
perm_pvals <- sapply(seq_len(length(tmp_dif)), function(i){
  # Get where on the distribution the observed value falls
  tmp_pval <- mean(perm_diff[i, ] >= tmp_dif[i])
  
  # Get the double tailed p value
  pval <- min(c(tmp_pval, 1 - tmp_pval))
  return(pval)
})

adj_perm_pvals <- p.adjust(p = perm_pvals, method = "bonferroni")
```

Join bootstrapped P values with proportions
```{r}
perm_df <- data.frame(t(adj_perm_pvals))
names(perm_df) <- names(prop_df)

hm_df_2 <- dplyr::bind_rows(prop_df, perm_df)
rownames(hm_df_2) <- c(rownames(prop_df), "pval")
```

Plot heatmap
```{r}
hm_tmp <- t(hm_df_2) %>%
  data.frame() %>% 
  tibble::rownames_to_column("cell_type") %>% 
  tidyr::pivot_longer(cols = c("Non.Tumoral", "Tumoral", "pval"),
                      names_to = "status",
                      values_to = "proportion") %>%
  dplyr::mutate(
    status = factor(status, levels = c("Non.Tumoral", "Tumoral", "pval"), labels = c("Non Tumoral", "Tumoral", "Adj. Pval"))
    ) %>%
  dplyr::left_join(pdac_plt_names, by = c("cell_type" = "df_name")) %>%
  dplyr::mutate(proportion_mod = dplyr::if_else(status == "Adj. Pval", 0, proportion),
                plt_name = as.character(plt_name),
                plt_name = dplyr::if_else(plt_name == "Macro_1", "Macrophage", plt_name),
                plt_name = dplyr::if_else(plt_name == "Mono", "Monocyte", plt_name),
                plt_name = stringr::str_wrap(string = plt_name, width = 20),
                pval = dplyr::if_else(status == "Adj. Pval", proportion, NA_real_)) %>% 
  ggplot(ggplot2::aes(x = status,
             y = plt_name, 
             fill= proportion_mod)) + 
  ggplot2::geom_tile(ggplot2::aes(alpha = proportion_mod)) +
  ggplot2::geom_text(ggplot2::aes(label = round(pval, 2)),
            show_guide  = FALSE,
            size = 7,
            family = "Arial") +
  ggplot2::theme_classic() +
  ggplot2::labs(title = "Cell type enrichment per region",
       y = "",
       x = "",
       fill = "Proportion",
       alpha = "Proportion") +
  ggplot2::scale_fill_gradient(low = "#FFF5F0",
                      high = "#99000D",
                      guide = "legend",
                      limits = c(0, 1),
                      breaks = seq(0, 1, 0.1)) +
  ggplot2::scale_alpha_continuous(guide = "legend",
                         limits = c(0, 1), 
                         breaks = seq(0, 1, 0.1)) +
  ggplot2::theme(
        axis.text.x = ggplot2::element_text(size = 20),
        axis.text.y = ggplot2::element_text(size = 20),
        legend.text = ggplot2::element_text(size = 20),
        legend.title = ggplot2::element_text(size = 22),
        plot.title = ggplot2::element_text(hjust = 0.5, face = "bold", size = 20),
        axis.title = ggplot2::element_text(size = 22),
        axis.line = ggplot2::element_blank())
hm_tmp
# saveRDS(object = hm_tmp,
#         file = sprintf("%s/%s/heatmap_pvalue_GSM3036911_status_2_territories.RDS",
#                                     an_pdac, plt_dir))

# saveRDS(object = tmp_plt_2,
#         file = sprintf("%s/%s/GSM3036911_status_2_territories.RDS",
#                                     an_pdac, plt_dir))

# cowplot::save_plot(plotlist = list(hm_tmp, tmp_plt_2),
#                  filename = sprintf("%s/%s/heatmap_pvalue_GSM3036911_status_2_territories.pdf",
#                                     an_pdac, plt_dir),
#                  ncol = 2,
#                  width = 16, 
#                  height = 9, 
#                  res = 600)

# ggpubr::ggexport(plotlist = list(hm_tmp, tmp_plt_2),
#                  filename = sprintf("%s/%s/heatmap_pvalue_GSM3036911_status_2_territories_arr.jpg",
#                                     an_pdac, plt_dir),
#                  ncol = 2,
#                  width = 6000, 
#                  height = 4000, 
#                  res = 400)

cowplot::save_plot(plot = hm_tmp,
                 filename = here::here(sprintf("%s/%s/heatmap_pvalue_GSM3036911_status_2_territories.pdf",
                                    an_pdac, plt_dir)),
                 # ncol = 2,
                 base_width = 10, 
                 base_height = 15)
```

#### Box plots
Preparing the dataset
```{r}
decon_mtrx_immune <- immune_decon_mtrx_ls[[1]][[2]]

decon_df_immune <- data.frame(decon_mtrx_immune)

decon_df_immune$status_2_territories <- st_se@meta.data[, "status_2_territories"]

decon_mtrx_long <- decon_df_immune %>% 
  dplyr::select(-c("res_ss")) %>% 
  dplyr::mutate(total = 1) %>%
  tidyr::pivot_longer(cols = -c(status_2_territories, total), 
                      names_to = "cell_type", 
                      values_to = "proportion") %>% 
  dplyr::left_join(pdac_plt_names, by = c("cell_type" = "df_name")) %>% 
  dplyr::mutate(plt_name = as.character(plt_name),
                plt_name = dplyr::if_else(plt_name == "Macro_1", "Macrophage", plt_name),
                plt_name = dplyr::if_else(plt_name == "Mono", "Monocyte", plt_name),
                plt_name = dplyr::if_else(plt_name == "CD8 tumor-reactive (exhausted)", "CD8 tumor-reactive", plt_name))

## Add an empty level to get symmetric facets
# ct = unique(decon_mtrx_long$plt_name)
# decon_mtrx_long$plt_name_new <- factor(decon_mtrx_long$plt_name, levels = c(ct[1:(length(ct) - 2)], "", ct[(length(ct) - 1):length(ct)]))


```

Get Bonferroni adjusted P-values
```{r}
# y max per group
y_pos <- decon_mtrx_long %>% 
  group_by(cell_type) %>% 
  summarise(y_max = max(proportion)) %>% 
  dplyr::pull(y_max)

# annotation table with adjusted pvals and y-position of the labels
anno_df <- ggpubr::compare_means(proportion ~ status_2_territories,
                                 method = "wilcox.test",
                                 group.by = "plt_name",
                                 data = decon_mtrx_long,
                                 p.adjust.method = "bonferroni") %>%
  mutate(y_pos = y_pos - 0.1 * y_pos, 
         p.adj.txt = paste("p =", p.adj))
```

Plot faceted boxplots, subseting cell types of interest
```{r}
ct_interest <- c("M2 TAMs", "Proliferative T cells", "Recently activated CD4 T cells",
                 "Plasma B cells", "Pre-exhausted CD8 T cells", "Proinflamatory TAMs",
                 "Cytotoxic CD8 T cells")
bplt <- decon_mtrx_long %>%
  left_join(anno_df, by = "plt_name") %>%
  filter(plt_name %in% ct_interest) %>%
  filter(p.adj < 0.05) %>% 
  ggpubr::ggboxplot(.,
                    x = "status_2_territories",
                    y = "proportion", 
                    color = "status_2_territories",
                    fill = "status_2_territories",
                    add = "jitter",
                    alpha = 0.6,
                    facet.by = "plt_name",
                    scales = "free",
                    # palette = "npg",
                    palette = c("#00A087FF", "#E64B35FF"),
                    nrow = 2,
                    outlier.shape = NA) +
  # ggplot2::facet_wrap(~plt_name) + 
  ggpubr::geom_signif(
    data = anno_df %>% filter(p.adj < 0.05 & plt_name %in% ct_interest), 
    aes(xmin = group1, 
        xmax = group2, 
        annotations = p.adj.txt, 
        y_position = y_pos * 1,
        size = 20,
        family = "Arial"),
    manual = TRUE)

bplt <- bplt +
  ggplot2::labs(title = "Spot composition comparison",
                x = "Tissue stratification",
                y = "Capture location proportion",
                fill = "Stratification",
                color = "Stratification") +
  ggplot2::theme(
    text = ggplot2::element_text(family = "Arial"),
    strip.text = ggplot2::element_text(size = 23, face = "bold"),
    axis.text =  ggplot2::element_text(size = 20),
    axis.title = ggplot2::element_text(size = 25, face = "bold"),
    strip.background.x = ggplot2::element_rect(colour = "transparent",
                                             fill = "transparent"),
    legend.position = "none",
    plot.title = ggplot2::element_text(face = "bold", hjust = 0.5, size = 30),
    legend.text = ggplot2::element_text(size = 12),
    legend.title = ggplot2::element_text(size = 15),
    legend.key.size = ggplot2::unit(3, "line"))

# saveRDS(object = bplt,
#         file = sprintf("%s/%s/boxplot_GSM3036911_status_2_territories_filt.RDS",
#                        an_pdac, plt_dir))

cowplot::save_plot(
  plot = bplt,
  filename = sprintf("%s/%s/boxplot_GSM3036911_status_2_territories_filt.pdf",
                     an_pdac, plt_dir),
  base_width = 16,
  base_height = 14)
```

Plot faceted boxplots All cell types
```{r}
bplt <- decon_mtrx_long %>%
  ggpubr::ggboxplot(., x = "status_2_territories", y = "proportion", 
                  color = "status_2_territories", fill = "status_2_territories",
                  add = "jitter", alpha = 0.6,
                  facet.by = "plt_name",
                  scales = "free",
                  # palette = "npg",
                  palette = c("#00A087FF", "#E64B35FF"),
                  outlier.shape = NA) +
  # ggplot2::facet_wrap(~plt_name) + 
  ggpubr::geom_signif(
    data = anno_df, 
    aes(xmin = group1, 
        xmax = group2, 
        annotations = p.adj.txt, 
        y_position = y_pos * 1),
    manual = TRUE)

bplt <- bplt +
  ggplot2::labs(title = "Spot composition comparison",
                x = "Tissue stratification",
                y = "Capture location proportion",
                fill = "Stratification",
                color = "Stratification") +
  ggplot2::theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text =  element_text(size = 12),
    axis.title = element_text(size = 15, face = "bold"),
    legend.position = "right",
    plot.title = element_text(face = "bold", hjust = 0.5, size = 20),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 15),
    legend.key.size = unit(3,"line"),
    # strip.text = element_text(face="bold", size=9),
    strip.background = element_rect(size = 0.3))
bplt
ggpubr::ggexport(plotlist = list(bplt),
                 filename = sprintf("%s/%s/boxplot_GSM3036911_status_2_territories.pdf",an_pdac, plt_dir),
                 width = 16,
                 height = 12,
                 res = 600)
```

#### Analyzing status_3_territories
Transfer assignment to Immune cell mapping
```{r}
decon_mtrx_immune <- immune_decon_mtrx_ls[[1]][[2]]

decon_mtrx_immune[decon_mtrx_immune > 0] <- 1

decon_df_immune <- data.frame(decon_mtrx_immune)

decon_df_immune$status_3_territories <- st_se@meta.data[, "status_3_territories"]
```

Territory location
```{r}
# Inverse Y to flip the coordinates
point_size <- "status_3_territories"
point_colour <- "status_3_territories"
data_df <- data.frame(st_se@meta.data)
data_df$y_inv <- 36 - data_df$y

# data_df[, point_size] <- if_else(data_df[, point_size] == 0, NA_real_, data_df[, point_size])

strat_3_plt <- data_df %>%
  dplyr::mutate(status_3_territories = dplyr::if_else(data_df$status_3_territories == "Removed manually",
                                                       "Excluded", status_3_territories)) %>% 
  ggplot2::ggplot(ggplot2::aes_string("x", "y_inv",
                                    color = "status_3_territories" 
                                    # alpha = point_alpha
                                    )) +
  ggplot2::xlim(1, 33) +
  ggplot2::ylim(1, 35) +
  # Layer 1 - Plot image
  ggplot2::annotation_custom(st_se@images[[1]],
                             xmin = 1,
                             xmax = 33,
                             ymin = 1,
                             ymax = 35) +
  # Layer 2 - Plot points
  geom_point(size = 5, alpha = 0.8) +
  labs(title = "Tissue stratification",
       color = "Stratification") +
  coord_fixed(1) +
  theme_void() +
  scale_color_manual(values = c("lightgrey", "#4DBBD5FF", "#00A087FF", "#E64B35FF")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 20))

strat_3_plt
saveRDS(object = strat_3_plt,
        file = sprintf("%s/%s/GSM3036911_status_3_territories.RDS",
                                    an_pdac, plt_dir))

cowplot::save_plot(
  plot = strat_3_plt +
    guides(color = guide_legend(override.aes = list(size = 10))) +
                   guides(fill = guide_legend(override.aes = list(size = 10))) +
                   theme(legend.title = element_text(size = 15, face = "bold"),
                         legend.text = element_text(size = 12)),
  filename = sprintf("%s/%s/GSM3036911_status_3_territories.pdf", 
                      an_pdac, plt_dir),
  base_width = 10,
  base_height = 8)
```

Proportion heatmap
```{r}
tmp_df <- decon_df_immune %>%
  dplyr::filter( status_3_territories != "Removed manually") %>% 
  dplyr::select(-c("res_ss")) %>% 
  dplyr::mutate(total = 1) %>% 
  dplyr::group_by(status_3_territories) %>% 
  dplyr::summarise_if(is.numeric, ~sum(.)) %>% 
  data.frame() %>% 
  tibble::column_to_rownames("status_3_territories")

prop_df <- tmp_df[, 1:(ncol(tmp_df) - 1)] / tmp_df[, ncol(tmp_df)]

# tmp_dif <- sapply(colnames(prop_df), function(i) prop_df[1, i] - prop_df[2, i])

```


```{r}
hm_tmp <- t(prop_df) %>%
  data.frame() %>% 
  tibble::rownames_to_column("cell_type") %>% 
  dplyr::filter(cell_type != "res_ss") %>% 
  tidyr::pivot_longer(cols = c("Non.Tumoral", "Intermediate", "Tumoral"), 
                      names_to = "health_status", 
                      values_to = "proportion") %>%
  dplyr::mutate(health_status = factor(health_status, levels = c("Non.Tumoral", "Intermediate", "Tumoral"))) %>%
  dplyr::left_join(pdac_plt_names, by = c("cell_type" = "df_name")) %>% 
  ggplot(aes(x = health_status, 
             y = plt_name, 
             fill= proportion)) + 
  geom_tile() +
  geom_text(aes(label = round(proportion, 2)), 
            show_guide  = FALSE) +
  labs(title = "GSM3036911 Cell-type enrichment by region",
       y = "Cell types", 
       x = "",
       fill = "Proportion") +
  theme_classic() +
  scale_fill_gradient(low = "#FFF5F0",
                      high = "#99000D",
                      guide = "legend",
                      limits = c(0, 1),
                      breaks = seq(0, 1, 0.1)) +
  scale_alpha_continuous(guide = "legend",
                         limits = c(0, 1), 
                         breaks = seq(0, 1, 0.1)) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 15),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 17),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 20),
        axis.title = element_text(size = 15),
        axis.line = element_blank())


hm_tmp

saveRDS(object = hm_tmp,
        file = sprintf("%s/%s/heatmap_GSM3036911_status_3_territories.RDS", 
                                      an_pdac, plt_dir))
ggpubr::ggarrange(plotlist = list(hm_tmp, strat_3_plt),
                  labels = c("A", "B")) %>% 
  ggpubr::ggexport(.,
                   filename = sprintf("%s/%s/heatmap_GSM3036911_status_3_territories.pdf", 
                                      an_pdac, plt_dir),
                   width = 16,
                   height = 9,
                   res = 600)
```

#### Box plots
Preparing the dataset
```{r}
decon_mtrx_immune <- immune_decon_mtrx_ls[[1]][[2]]

decon_df_immune <- data.frame(decon_mtrx_immune)

decon_df_immune$status_3_territories <- st_se@meta.data[, "status_3_territories"]

decon_mtrx_long <- decon_df_immune %>% 
  dplyr::filter(status_3_territories != "Removed manually") %>% 
  dplyr::select(-c("res_ss")) %>% 
  dplyr::mutate(total = 1) %>%
  tidyr::pivot_longer(cols = -c(status_3_territories, total), 
                      names_to = "cell_type", 
                      values_to = "proportion") %>% 
  dplyr::left_join(pdac_plt_names, by = c("cell_type" = "df_name")) %>% 
  dplyr::mutate(plt_name = as.character(plt_name))
```

Perform ANOVA and pairwise t-test and adjust P-values
```{r}
# Get max position for each facet
y_pos <- decon_mtrx_long %>%
  group_by(plt_name) %>%
  summarise(y.position = max(proportion))

# Define list of pairwise interactions to test
my_comparisons <- list( c("Intermediate", "Non-Tumoral"),
                        c("Non-Tumoral", "Tumoral"),
                        c("Intermediate", "Tumoral") )


# Annova test annotation table with adjusted pvals and y-position of the labels
anno_df_anova <- ggpubr::compare_means(proportion ~ status_3_territories,
                                 method = "anova",
                                 group.by = "plt_name",
                                 data = decon_mtrx_long,
                                 p.adjust.method = "bonferroni",
                                 comparisons = my_comparisons) %>%
  dplyr::left_join(pdac_plt_names, by = "plt_name") %>% 
  dplyr::left_join(y_pos, by = "plt_name")


# T-test annotation table with adjusted pvals and y-position of the labels
anno_df_wilcox <- ggpubr::compare_means(proportion ~ status_3_territories,
                                 method = "wilcox.test",
                                 group.by = "plt_name",
                                 data = decon_mtrx_long,
                                 p.adjust.method = "bonferroni",
                                 comparisons = my_comparisons) %>%
  dplyr::left_join(pdac_plt_names, by = "plt_name") %>% 
  dplyr::left_join(y_pos, by = "plt_name") %>% 
  dplyr::mutate(y.position = if_else(group1 == "Non-Tumoral" & 
                                       group2 == "Intermediate", 
                                     y.position * 0.7, 
                                     if_else(group1 == "Intermediate" &
                                               group2 == "Tumoral", 
                                             y.position * 0.8,
                                             y.position * 0.93)))

```

```{r}
p <- ggpubr::ggboxplot(decon_mtrx_long,
                       x = "status_3_territories",
                       y = "proportion",
                       color = "status_3_territories",
                       fill = "status_3_territories",
                       add = "jitter",
                       alpha = 0.6,
                       facet.by = "plt_name",
                       scales = "free",
                       repel = TRUE,
                       palette = c("#00A087FF", "#4DBBD5FF", "#E64B35FF"),
                       outlier.shape = NA)

bplt <- p + 
  ggpubr::stat_pvalue_manual(anno_df_wilcox, label = "p = {p.adj}", size = 6) +
  ggplot2::labs(
    # title = "PDAC-A Immune capture location composition comparison 3 section stratification",
                x = "Tissue stratification",
                y = "Capture location proportion",
                fill = "Stratification",
                color = "Stratification") +
  ggplot2::theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text =  element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    legend.position = "right",
    plot.title = element_text(face = "bold", hjust = 0.5, size = 25),
    legend.text = element_text(size = 17),
    legend.title = element_text(size = 20),
    legend.key.size = unit(3,"line"))


saveRDS(object = bplt,
        file = sprintf("%s/%s/boxplot_GSM3036911_status_3_territories.RDS",
                       an_pdac, plt_dir))

cowplot::save_plot(
  plot = bplt,
  filename = sprintf("%s/%s/boxplot_GSM3036911_status_3_territories.pdf",
                     an_pdac, plt_dir),
  base_width = 32,
  base_height = 24)
```

Boxplot of cells of interest
```{r}
ct_interest <- c("M2 TAMs", "Proliferative T cells",
                 "Recently activated CD4 T cells", "Plasma B cells",
                 "Pre-exhausted CD8 T cells", "Proinflamatory TAMs")

p <- decon_mtrx_long %>%
  dplyr::filter(plt_name %in% ct_interest) %>% 
  ggpubr::ggboxplot(.,
                       x = "status_3_territories",
                       y = "proportion",
                       color = "status_3_territories",
                       fill = "status_3_territories",
                       add = "jitter",
                       alpha = 0.6,
                       facet.by = "plt_name",
                       scales = "free",
                       repel = TRUE,
                       palette = c("#00A087FF", "#4DBBD5FF", "#E64B35FF"),
                       outlier.shape = NA)

anno_df_wilcox_sub <- anno_df_wilcox %>% filter(plt_name %in% ct_interest)

bplt_int <- p + 
  ggpubr::stat_pvalue_manual(anno_df_wilcox_sub, label = "p = {p.adj}", size = 6) +
  ggplot2::labs(
    # title = "PDAC-A Immune capture location composition comparison 3 section stratification",
                x = "Tissue stratification",
                y = "Capture location proportion",
                fill = "Stratification",
                color = "Stratification") +
  ggplot2::theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text =  element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    legend.position = "right",
    plot.title = element_text(face = "bold", hjust = 0.5, size = 25),
    legend.text = element_text(size = 17),
    legend.title = element_text(size = 20),
    legend.key.size = unit(3,"line"))

```


##### Arrange 3 territory stratification
```{r}
cowplot::plot_grid(strat_3_plt + theme(plot.title = element_text(size = 20)),
                   bplt_int,
                   align = "h",
                   nrow = 1,
                   ncol = 2,
                   rel_widths = c(1, 1.5)) %>% 
  cowplot::save_plot(filename = sprintf("%s/%s/3_strat_arrange.pdf",
                                        an_pdac, plt_dir),
                     base_width = 25,
                     base_height = 10)
```

