---
title: "Allen 70k  plots"
author: "Marc Elosua-Bayes"
date: "4/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This is an Rmarkdown document where we are going to plot the results from the deconvolution.
We are going to plot:
* Spatial location + gene expression + ISH data
* Scatterpie of all cell types + only spots with cell types of interest

## Libraries
```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(purrr)
library(SPOTlight)
library(NMF)
library(nnls)
library(cowplot)
library(imager)
library(scatterpie)
```

## Paths
```{r}
tech <- "sc"
tissue <- "allen_ref_70k"
dwn_smplng <- "both"
org <- "mm"
source("misc/paths_vrs.R")
source("utils/bin.r")
```

## Set common parameters
```{r}
clust_vr <- "subclass_label"
cl_n <- 100
method <- "nsNMF"
transf <- "uv"
hvg <- 3000
FC <- 1
pct1 <- 0.9
seed_id <- 123

id_nmf <- sprintf("cln-%s_transf-%s_method-%s_hvg-%s_FC-%s_pct1-%s_seed-%s", 
                  cl_n, transf, method, hvg, FC, pct1, seed_id)
data_dir <- "data/MusMusculus/allen_reference"
options(stringsAsFactors = FALSE)


library(RColorBrewer)
n <- 60
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector  <-  unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
col_low <- "green"; col_high <- "red"
```

## Load data
```{r}
brain <- readRDS(sprintf("%s/%s/brain1_processed.RDS", an_mouse, robj_dir))
decon_mtrx <- readRDS(file = sprintf("decon_mtrx_%s_%s.RDS", id_comp, id_nmf))
```

Join the data so its easier to work with
```{r}
cell_types <- colnames(decon_mtrx)[which(colnames(decon_mtrx) != "res_ss")]

# decon_mtrx_prop <- round(decon_mtrx[, cell_types] / rowSums(decon_mtrx[, cell_types]), 4)
# decon_mtrx_prop[is.na(decon_mtrx_prop)] <- 0
brain@meta.data <- cbind(brain@meta.data, decon_mtrx_prop)
cell_types_metadata <- colnames(brain@meta.data)[colnames(brain@meta.data) %in% cell_types]
```

To maintain consistent colors for each cell type between different plots we will create a dataframe of equivalencies which we will draw colors from
```{r}
col_df <- data.frame(cell_types = sort(c(cell_types, "L2.3")),
           col_vector = col_vector[1:(length(cell_types)+1)])
```

To return the right name in the plots:
```{r}
# ct_names <- c("Astro", "CA1sp", "CA1sp.SUB-sp.Kcnip1", "CA2sp.IG", "CA3sp", 
#               "Car3", "CR", "DG", "Doublet", "Endo", "IT.RHP.Dcn", 
#               "L2.3.IT.Cdc14a", "L2.3.IT.Cxcl14", "L2.3.IT.Ndst4.Endou", 
#               "L2.3.IT.Otof", "L2.3.IT.Plch1", "L4.IT", "L4.5.IT", "L5.ET", 
#               "L5.IT", "L6.CT" ,"L6.IT", "L6b", "Lamp5", "Lamp5.Lhx6", 
#               "Low.Quality", "Ly6g6e", "Macrophage", "Meis2", "NP", "Oligo",
#               "PIR.Six3", "POST-PRE-PAR.Ptgfr", "Pvalb", "RHP.Cplx3", 
#               "RSP_ACA.IT.Scnn1a", "RSP.ACA.L4.5.IT", "Serpinf1", "SMC", "Sncg",
#               "Sncg.Ndnf.HPF", "Sst", "Sst.Chodl", "SUB-Sp.Ndst4", "Unknown", 
#               "Vip", "VLMC", "L2.3")

plt_names <- c("Astrocytes", "CA1sp", "CA1sp/SUB-sp Kcnip1", "CA2sp/IG", "CA3sp",
               "Car3", "Cajal-Retzius", "DG", "Doublet", "Endothelium", 
               "IT RHP Dcn", "L2/3 IT Cdc14a", "L2/3 IT Cxcl14", 
               "L2/3 IT Ndst4 Endou", "L2/3 IT Otof", "L2/3 IT Plch1", "L4 IT", 
               "L4/5 IT", "L5 ET", "L5 IT", "L6 CT" ,"L6 IT", "L6b", "Lamp5", 
               "Lamp5 Lhx6", "Low quality", "Ly6g6e", "Macrophage", "Meis2", 
               "Neural Progenitor", "Oligodendrocyte", "PIR Six3", 
               "POST-PRE-PAR Ptgfr", "Pvalb", "RHP Cplx3", "RSP/ACA IT Scnn1a", 
               "RSP/ACA L4/5 IT", "Serpinf1", "Smooth Muscle Cells", "Sncg", 
               "Sncg/Ndnf HPF", "Sst", "Sst Chodl", "SUB-Sp Ndst4", "Unknown", 
               "Vip", "VLMC", "L2/3")

ct_df <- data.frame(ct_names = gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
                                            x = plt_names,
                                            perl = TRUE), 
                    plt_names = plt_names)

```

## Functions
Function to plot both anterior and posterior scatterpie slides together
```{r}
joint_scatterpie_fun <- function(se_obj, cell_types_all, img_path1, img_path2, 
                                 slice1, slice2, cell_types_interest, 
                                 return_legend, img_alpha, scatterpie_alpha, 
                                 pie_scale, col_df, arr_leg = TRUE) {
  
  anterior_plt <- spatial_scatterpie(se_obj = se_obj,
                   cell_types_all = cell_types_all,
                   img_path = img_path1,
                   slice = slice1, 
                   cell_types_interest = cell_types_interest, 
                   return_legend = return_legend,
                   img_alpha = img_alpha,
                   scatterpie_alpha = scatterpie_alpha,
                   pie_scale = pie_scale,
                   col_df = col_df)
  posterior_plt <- spatial_scatterpie(se_obj = se_obj,
                     cell_types_all = cell_types_all,
                     img_path = img_path2,
                     slice = slice2, 
                     cell_types_interest = cell_types_interest,
                     return_legend = return_legend,
                     img_alpha = img_alpha,
                     scatterpie_alpha = scatterpie_alpha,
                     pie_scale = pie_scale,
                     col_df = col_df)
  
  tmp_plt <- scatterpie_plot(se_obj = se_obj,
                 cell_types_all = cell_types_all,
                 slice = slice1,
                 scatterpie_alpha = scatterpie_alpha, 
                 col_df = col_df) +
             # labs(fill = "") +
             theme(legend.position = "bottom",
                   legend.text = element_text(colour = "#3b3a39", size = 9),
                   legend.title = element_text(colour = "#3b3a39", vjust = 1, hjust = 0.5),
                   legend.key.size = unit(1, "line")) +
             guides(fill = guide_legend(nrow = 5, override.aes = list(size = 0.1)))
  
  legend_gtable <- get_legend(tmp_plt)
  
  scatterpie_tmp <- cowplot::plot_grid(plotlist = list(anterior_plt[[1]],
                                              posterior_plt[[1]]),
            ncol = 2,
            nrow = 1, 
            rel_widths = c(1, 1))
  
  if (arr_leg) {
    scatterpie_tmp <- cowplot::plot_grid(plotlist = list(scatterpie_tmp,
                                              legend_gtable),
            ncol = 1,
            nrow = 2, 
            rel_heights = c(1, 0.15))
  } 
  
  
  return(scatterpie_tmp)
}

join_spatial_plots <- function(spatial_obj, ct) {
  ###########################################################################################################
  #### This functions takes in a feature to plot in 2 spatial slides and returns it with a unique legend ####
  ###########################################################################################################
  library(ggplot2)
  library(ggpubr)
  library(cowplot)
  library(Seurat)
  
  # Set common legend theme
  legend_theme <-  theme(legend.text = element_text(colour = "#3b3a39", size = 10),
                       legend.title = element_text(colour = "#3b3a39", vjust = 1))
  
  orig_pt <- SpatialFeaturePlot(spatial_obj,
                   features = ct,
                   pt.size.factor = 1,
                   alpha = c(0, 1)) + 
    theme(plot.title = element_text(hjust = 0.5))
  
  ### Modify individual plots ###
  orig_pt_ant <- orig_pt[[1]] +
    cowplot::theme_map() + 
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))+ 
    scale_fill_gradientn(
      # colours = colorRamps::matlab.like2(10),
      colours = heat.colors(10, rev = TRUE),
      limits = c(0, 1)) +
    # Set aplha limit as if the upper limit was 1 so that the alpha proportions are plot independent
    scale_alpha_continuous(range = c(0, max(orig_pt[[1]][["data"]][[ct]]))) +
    labs(title = "", 
       fill = "Proportion")
  
  orig_pt_post <- orig_pt[[2]] +
    labs(title = "", fill = ct) +
    cowplot::theme_map() + 
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) + 
    scale_fill_gradientn(
      # colours = colorRamps::matlab.like2(10),
      colours = heat.colors(10, rev = TRUE),
      limits = c(0, 1)) +
    # Set aplha limit as if the upper limit was 1 so that the alpha proportions are plot independent
    scale_alpha_continuous(range = c(0, max(orig_pt[[2]][["data"]][[ct]])))

  return(list(orig_pt_ant, orig_pt_post))
}

```

## Plots
### Compositions
In this sections we are going to build compositions of a cell's spatial location + marker gene for that cell type in the tissue + ISH image of that cell type

#### Cell types of interest

##### Hipoccampus cells
Start by Making a dictionary of: cell type name in the data - cell type name in the plot - marker gene
```{r}
ct_dict <- list()
ct_dict[["CA1sp"]][["gene"]] <- "Fibcd1"
ct_dict[["CA1sp"]][["name"]] <- "CA1"
ct_dict[["CA1sp"]][["plot_name"]] <- "Cornu Ammonis 1 sp"

ct_dict[["CA2sp.IG"]][["gene"]] <- "Ccdc3"
ct_dict[["CA2sp.IG"]][["name"]] <- "CA2"
ct_dict[["CA2sp.IG"]][["plot_name"]] <- "Cornu Ammonis 2 sp"

ct_dict[["CA3sp"]][["gene"]] <- "Pvrl3"
ct_dict[["CA3sp"]][["name"]] <- "CA3"
ct_dict[["CA3sp"]][["plot_name"]] <- "Cornu Ammonis 3 sp"

ct_dict[["DG"]][["gene"]] <- "Prox1"
ct_dict[["DG"]][["name"]] <- "DG"
ct_dict[["DG"]][["plot_name"]] <- "Dentate Gyrus"
```

```{r}
# Set common legend theme
legend_theme <-  theme(legend.text = element_text(colour = "#3b3a39", size = 8),
                     legend.title = element_text(colour = "#3b3a39", vjust = 1))

join_spatial_plots_2 <- function(spatial_obj, feat) {
  ###########################################################################################################
  #### This functions takes in a feature to plot in 2 spatial slides and returns it with a unique legend ####
  ###########################################################################################################
  library(ggplot2)
  library(ggpubr)
  library(cowplot)
  library(Seurat)
  
  # Set common legend theme
  legend_theme <-  theme(legend.text = element_text(colour = "#3b3a39", size = 10),
                       legend.title = element_text(colour = "#3b3a39", vjust = 1))
  
  orig_pt <- SpatialFeaturePlot(spatial_obj,
                   features = feat,
                   pt.size.factor = 1,
                   alpha = c(0, 1)) + 
    theme(plot.title = element_text(hjust = 0.5))
  
  ### Modify individual plots ###
  orig_pt_ant <- orig_pt[[1]] +
    cowplot::theme_map() + 
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
  
  # Check if its a cell type, if it is set limits 0-1
  if (feat %in% colnames(spatial_obj@meta.data)) {
    orig_pt_ant <- orig_pt_ant + 
      scale_fill_gradientn(
        colours = heat.colors(10, rev = TRUE),
        limits = c(0, 1)) +
      labs(title = "", 
         fill = "Proportion")

  } else {
    # if its gene expression set common limit for both plots
    up_lim <- max(spatial_obj@assays$SCT@data[rownames(spatial_obj@assays$SCT@data) == feat, ])    
    
    if(sum(spatial_obj@assays$SCT@data[feat, spatial_obj$slice == "Anterior"]) == 0) {
      orig_pt_ant <- orig_pt_ant +
          # geom_point(alpha = 0) +
          scale_fill_gradientn(
            colours = NA,
            limits = c(0, up_lim)) +
          scale_alpha(range = c(0, 0)) +
          labs(title = "", 
             fill = sprintf("%s\nExpression", feat))
    } else {
      orig_pt_ant <- orig_pt_ant + 
          scale_fill_gradientn(
            colours = colorRamps::matlab.like2(10),
              limits = c(0, up_lim)) +
          labs(title = "", 
             fill = sprintf("%s\nExpression", feat))

    }
      }
  
  orig_pt_post <- orig_pt[[2]] +
    labs(title = "", fill = nam) +
    cowplot::theme_map() + 
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
  
  # Check if its a cell type, if it is set limits 0-1
  if (feat %in% colnames(spatial_obj@meta.data)) {
    orig_pt_post <- orig_pt_post + 
      scale_fill_gradientn(
        colours = heat.colors(10, rev = TRUE),
        limits = c(0, 1))
  } else {
    up_lim <- max(spatial_obj@assays$SCT@data[rownames(spatial_obj@assays$SCT@data) == feat, ])
    
    if(sum(spatial_obj@assays$SCT@data[feat, spatial_obj$slice == "Posterior"]) == 0) {
      orig_pt_post <- orig_pt_post +
        geom_point(alpha = 0) +
      scale_fill_gradientn(
        colours = NA,
        limits = c(0, up_lim)) +
        scale_alpha(range = c(0, 0))
    } else {
      orig_pt_post <- orig_pt_post + 
      scale_fill_gradientn(
        colours = colorRamps::matlab.like2(10),
        limits = c(0, up_lim))
    }
    
  }

  return(list(orig_pt_ant, orig_pt_post))

}
```

We will use imager to work with images: importing + formating + plotting.
```{r}
ct <- "CA2sp.IG"
ct_ls <- c("CA1sp", "CA2sp.IG", "CA3sp", "DG")
plot_ls <- lapply(ct_ls, function(ct) {
  feat <- ct_dict[[ct]][["gene"]] 
  nam <- ct_dict[[ct]][["name"]]
  
  # If a gene doesn't exist add its row as all 0
  if (!feat %in% rownames(brain@assays$SCT@data)) {
    feat_mtrx <- Matrix::Matrix(0, nrow = 1, ncol = ncol(brain@assays$SCT@data), sparse = TRUE)
    rownames(feat_mtrx) <- feat
    brain@assays$SCT@var.features <- c(brain@assays$SCT@var.features, feat)
    brain@assays$SCT@data <- rbind(brain@assays$SCT@data, feat_mtrx)
  }
  # Plot cell type
  suppressMessages(ct_arr <- join_spatial_plots_2(spatial_obj = brain, feat = ct))
  legend_ct <- get_legend(ct_arr[[1]] + legend_theme)
  
  # Plot gene marker for that cell type
  suppressMessages(feat_arr <- join_spatial_plots_2(spatial_obj = brain, feat = feat))
  legend_feat <- get_legend(feat_arr[[1]] + legend_theme)
  
  ### Load image to plot ###
  fn_all <- list.files("analysis/mouse_brain/img")
  fn_ls <- fn_all[grepl(pattern = paste0("^", nam), x = fn_all)]
  
  fn_anterior <- fn_ls[grepl(pattern = "anterior_crop", x = fn_ls)]
  anterior_ish <- plot_image(img_path = sprintf("analysis/mouse_brain/img/%s", fn_anterior)) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    coord_fixed(ratio = 0.925, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on")
  
  fn_posterior <- fn_ls[grepl(pattern = "posterior_crop", x = fn_ls)]
  posterior_ish <- plot_image(img_path = sprintf("analysis/mouse_brain/img/%s", fn_posterior)) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    coord_fixed(ratio = 0.925, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on")
  
  ct_plt <- cowplot::plot_grid(plotlist = list(ct_arr[[1]] + theme(legend.position = "none"),
                                               ct_arr[[2]] + theme(legend.position = "none")),
                               align = "vh",
                               # axis = "trbl",
                               ncol = 2,
                               nrow = 1)
  
  feat_plt <- cowplot::plot_grid(plotlist = list(feat_arr[[1]] + theme(legend.position = "none"),
                                               feat_arr[[2]] + theme(legend.position = "none")),
                               align = "vh",
                               # axis = "trbl",
                               ncol = 2,
                               nrow = 1)
  ish_plt <- cowplot::plot_grid(plotlist = list(anterior_ish,
                                               posterior_ish),
                               align = "vh",
                               # axis = "trbl",
                               ncol = 2,
                               nrow = 1)
  
  plt_arr <- cowplot::plot_grid(plotlist = list(ct_plt,
                                                feat_plt,
                                                NULL,
                                                ish_plt),
                                nrow = 4, 
                                ncol = 1, 
                                # align = "vh", 
                                axis = "tblr",
                                rel_heights = c(1, 1, 0.075, 1))
  
  leg_arr <- cowplot::plot_grid(plotlist = list(legend_ct, 
                                                legend_feat, 
                                                NULL),
                                nrow = 3, 
                                ncol = 1)
  
  final_arr <- cowplot::plot_grid(plotlist = list(plt_arr, leg_arr), 
                                  nrow = 1, 
                                  ncol = 2, 
                                  rel_widths = c(1 ,0.2), 
                                  axis = "trbl") %>% 
    ggpubr::annotate_figure(p = .,
                            top = ggpubr::text_grob(sprintf("%s", ct_dict[[ct]][["plot_name"]]), 
                                                  face = "bold", 
                                                  size = 20, vjust = 1))
  
  
  #### Save plot ####
  ggpubr::ggexport(plotlist = list(final_arr),
                         filename = sprintf("%s/%s/Fig_%s_arrange.jpeg", an_mouse, plt_dir, nam), 
                         width = 3000, 
                         height = 3700, 
                         res = 300)

  return(final_arr + theme(plot.background = element_rect(fill = NA, colour = "black", size = 2)))
})

plot_ls
```

##### Join Hippocampus plots
```{r}
hipp_plt <- cowplot::plot_grid(plotlist = list(plot_ls[[1]], NULL, plot_ls[[2]],
                                   NULL, NULL, NULL,
                                   plot_ls[[3]], NULL, plot_ls[[4]]), 
                   nrow = 3,
                   ncol = 3,
                   # align = "hv", 
                   rel_widths = c(1, 0.1, 1),
                   rel_heights = c(1, 0.1, 1),
                   axis = "trbl") + 
            theme(plot.margin = margin(1, 1, 1, 1, "cm"),
                  plot.background = element_rect(fill = "White"),
                  panel.background = element_blank())

cowplot::save_plot(plot = hipp_plt,
                     filename =
                   sprintf("%s/%s/Figure_ZZZ_hippocampus_arrange_%s_%s.pdf",
                           an_mouse, plt_dir, id_comp, id_nmf), 
                     base_width = 15, 
                     base_height = 19)

ggpubr::ggexport(plotlist = list(hipp_plt),
                     filename =
                   sprintf("%s/%s/Figure_ZZZ_hippocampus_arrange_%s_%s.jpeg",
                           an_mouse, plt_dir, id_comp, id_nmf), 
                     width = 5000,
                     height = 6500,
                     res = 300)

```


#### Scatterpie spatial plots
##### All cell types
```{r}
all_ct_plt <- joint_scatterpie_fun(se_obj = brain, 
                     cell_types_all = cell_types_metadata,
                     img_path1 = "data/MusMusculus/sag_ant_1/spatial/tissue_lowres_image.png",
                     img_path2 = "data/MusMusculus/sag_post_1/spatial/tissue_lowres_image.png",
                     slice1 = "Anterior",
                     slice2 = "Posterior",
                     cell_types_interest = NULL,
                     return_legend = TRUE,
                     img_alpha = 1,
                     scatterpie_alpha = 1,
                     pie_scale = 0.2,
                     col_df = col_df)   

saveRDS(object = all_ct_plt,
        file = sprintf("%s/%s/Figure_YYY_spatial_scatterpie_allen_%s_%s.RDS",
                                an_mouse, plt_dir, id_comp, id_nmf))

ggpubr::ggexport(plotlist = list(all_ct_plt), 
               filename = sprintf("%s/%s/Figure_YYY_spatial_scatterpie_allen_%s_%s.pdf",
                                an_mouse, plt_dir, id_comp, id_nmf), 
               width = 16, 
               height = 12, 
               res = 600)

ggpubr::ggexport(plotlist = list(all_ct_plt + theme(legend.position = "none")), 
               filename = sprintf("%s/%s/Figure_YYY_spatial_scatterpie_allen_%s_%s.jpeg",
                                an_mouse, plt_dir, id_comp, id_nmf), 
               width = 6000, 
               height = 4000, 
               res = 400)

```

##### Brain Layer cells
Next we want to plot neuronal layers and subtypes in both anterior and posterior sections
```{r eval = FALSE}
cortex_layers <- c("L2.3.IT.Cdc14a", "L2.3.IT.Cxcl14", "L2.3.IT.Ndst4.Endou", "L2.3.IT.Otof", "L2.3.IT.Plch1", "L4.IT", "L4.5.IT", "L5.ET", "L5.IT", "L6.CT", "L6.IT", "L6b")

joint_scatterpie_fun(se_obj = brain_mod,
                     cell_types_all = cell_types_metadata,
                     img_path1 = "data/MusMusculus/sag_ant_1/spatial/tissue_lowres_image.png",
                     img_path2 = "data/MusMusculus/sag_post_1/spatial/tissue_lowres_image.png",
                     slice1 = "Anterior",
                     slice2 = "Posterior",
                     cell_types_interest = cortex_layers,
                     return_legend = TRUE,
                     img_alpha = 1,
                     scatterpie_alpha = 1,
                     pie_scale = 0.2,
                     col_df = col_df) %>%
  ggpubr::ggexport(plotlist = list(.), 
                   filename = sprintf("%s/cortex_spatial_scatterpie_allen_%s_%s.pdf",
                                    an_mouse, id_comp, id_nmf), 
                   width = 12, 
                   height = 9, 
                   res = 600)

```

##### General cortex
Combine all L2/3 neurons
```{r}
brain_mod <- brain
brain_mod@meta.data[["L2.3"]] <- rowSums(brain@meta.data[, c("L2.3.IT.Cdc14a", "L2.3.IT.Cxcl14", "L2.3.IT.Ndst4.Endou", "L2.3.IT.Otof", "L2.3.IT.Plch1")])
brain_mod[["L2.3.IT.Cdc14a"]] <- NULL
brain_mod[["L2.3.IT.Cxcl14"]] <- NULL
brain_mod[["L2.3.IT.Ndst4.Endou"]] <- NULL
brain_mod[["L2.3.IT.Otof"]] <- NULL
brain_mod[["L2.3.IT.Plch1"]] <- NULL
```

Scatterpie plot with all the celltypes together
```{r}
cortex_layers_mod <- c("L2.3", "L4.IT", "L4.5.IT", "L5.ET", "L5.IT", "L6.CT", "L6.IT", "L6b")
cell_types_metadata_mod <- c(cell_types_metadata[!cell_types_metadata %in% c("L2.3.IT.Cdc14a", "L2.3.IT.Cxcl14", "L2.3.IT.Ndst4.Endou", "L2.3.IT.Otof", "L2.3.IT.Plch1")], "L2.3")

# general_cortex_plt <- joint_scatterpie_fun(se_obj = brain_mod,
#                      cell_types_all = cell_types_metadata_mod,
#                      img_path1 = "data/MusMusculus/sag_ant_1/spatial/tissue_lowres_image.png",
#                      img_path2 = "data/MusMusculus/sag_post_1/spatial/tissue_lowres_image.png",
#                      slice1 = "Anterior",
#                      slice2 = "Posterior",
#                      cell_types_interest = cortex_layers_mod,
#                      return_legend = TRUE,
#                      img_alpha = 1,
#                      scatterpie_alpha = 1,
#                      pie_scale = 0.2,
#                      col_df = col_df)

anterior_plt <- spatial_scatterpie(se_obj = brain_mod,
                 cell_types_all = cell_types_metadata_mod,
                 img_path = "data/MusMusculus/sag_ant_1/spatial/tissue_lowres_image.png",
                 slice = "Anterior", 
                 cell_types_interest = cortex_layers_mod, 
                 return_legend = TRUE,
                 img_alpha = 1,
                 scatterpie_alpha = 1,
                 pie_scale = 1,
                 col_df = col_df)

posterior_plt <- spatial_scatterpie(se_obj = brain_mod,
                   cell_types_all = cell_types_metadata_mod,
                   img_path = "data/MusMusculus/sag_post_1/spatial/tissue_lowres_image.png",
                   slice = "Posterior", 
                   cell_types_interest = cortex_layers_mod,
                   return_legend = TRUE,
                   img_alpha = 1,
                   scatterpie_alpha = 1,
                   pie_scale = 1,
                   col_df = col_df)

tmp_plt <- scatterpie_plot(se_obj = brain_mod,
               cell_types_all = cell_types_metadata_mod,
               slice = "Anterior",
               scatterpie_alpha = 1, 
               col_df = col_df) +
           labs(fill = "") +
           theme(legend.position = "bottom",
                 legend.text = element_text(colour = "#3b3a39", size = 9),
                 legend.title = element_text(colour = "#3b3a39", vjust = 1, hjust = 0.5),
                 legend.key.size = unit(1, "line"), legend.margin = margin(0,0,0,0)) +
           guides(fill = guide_legend(nrow = 5, override.aes = list(size = 0.1)))

legend_gtable <- get_legend(tmp_plt)

scatterpie_tmp <- cowplot::plot_grid(plotlist = list(anterior_plt[[1]] + theme(plot.margin = margin(0,0,0,0, "cm")),
                                            posterior_plt[[1]] + theme(plot.margin = margin(0,0,0,0, "cm"))),
          ncol = 2,
          nrow = 1)


general_cortex_plt <- cowplot::plot_grid(plotlist = list(scatterpie_tmp,
                                          legend_gtable),
        ncol = 1,
        nrow = 2, 
        rel_heights = c(1, 0.1)) +
  coord_fixed() +
  theme(plot.background = element_rect(fill = NA, colour = "black"))


saveRDS(object = general_cortex_plt,
        file = sprintf("%s/general_cortex_spatial_scatterpie_allen_%s_%s.pdf",
                                    an_mouse, id_comp, id_nmf))

ggpubr::ggexport(plotlist = list(general_cortex_plt), 
                   filename = sprintf("%s/general_cortex_spatial_scatterpie_allen_%s_%s.pdf",
                                    an_mouse, id_comp, id_nmf), 
                   width = 12, 
                   height = 9, 
                   res = 600)

ggpubr::ggexport(plotlist = list(general_cortex_plt + theme_void()), 
                   filename = sprintf("%s/%s/general_cortex_spatial_scatterpie_allen_%s_%s.jpeg",
                                    an_mouse, plt_dir, id_comp, id_nmf), 
                   width = 6000, 
                   height = 4000, 
                   res = 400)

```

Individual proportion plots
```{r eval = FALSE}
cortex_arrange <- lapply(cortex_layers_mod, function(ct) {
  # print(ct)
  ct_ls <- join_spatial_plots(spatial_obj = brain_mod, ct = ct)
  plt_tmp <- cowplot::plot_grid(plotlist = list(ct_ls[[1]] + theme(legend.position = "none"), 
                                                ct_ls[[2]] + theme(legend.position = "none")), 
                                nrow = 1, 
                                ncol = 2, 
                                align = "hv", 
                                labels = ct_df[ct_df$ct_names == ct, "plt_names"])

  # leg_grobtable <- get_legend(ct_ls[[1]])
  # plt_titleless <- cowplot::plot_grid(plotlist = list(plt_tmp, leg_grobtable), 
  #                    ncol = 2, 
  #                    nrow = 1, rel_widths = c(1, 0.2)) +
  #   theme(plot.background = element_rect(fill = NA, colour ="black", size = 2))
  
  return(plt_tmp)
})

saveRDS(object = cortex_arrange, 
        file = sprintf("%s/%s/general_cortex_ct_location.RDS",
                                    an_mouse, plt_dir, id_comp, id_nmf))

```

###### Arrangement
```{r eval = FALSE}
scatterpie_def <- round_corners_grob(scatterpie_tmp + theme(plot.background = element_rect(fill = NA, colour = "black", size = 2)))
gen_cort_arr <- cowplot::plot_grid(plotlist = list(cortex_arrange[[1]], NULL, cortex_arrange[[2]], NULL, cortex_arrange[[3]],
                                   NULL, NULL, NULL, NULL, NULL,
                                   cortex_arrange[[4]], NULL, scatterpie_def, NULL, cortex_arrange[[5]],
                                   NULL, NULL, NULL, NULL, NULL,
                                   cortex_arrange[[6]], NULL, cortex_arrange[[7]], NULL, cortex_arrange[[8]]),
                   nrow = 5,
                   ncol = 5,
                   rel_widths = c(1, 0.05, 1, 0.05, 1, 0.05),
                   rel_heights = c(1, 0.05, 1, 0.05, 1, 0.05),
                   labels = c("A", "", "B", "", "C",
                              rep("", 5), 
                              "D", "", "E", "", "F",
                              rep("", 5), 
                              "G", "", "H", "", "I"),
                   label_x = 0.025,label_y = 0.98)

ggpubr::ggexport(plotlist = list(gen_cort_arr),
                 filename = sprintf("%s/%s/Figure_QQQ_general_cortex_%s_%s.pdf", 
                                    an_mouse, plt_dir, id_comp, id_nmf),
                 width = 18,
                 height = 12,
                 res = 600)
```


##### Hippocapmus
```{r}
hip_layers <- c("CA1sp", "CA1sp.SUB-sp.Kcnip1", "CA2sp.IG", "CA3sp", "DG")

joint_scatterpie_fun(se_obj = brain, 
                     cell_types_all = cell_types_metadata,
                     img_path1 = "data/MusMusculus/sag_ant_1/spatial/tissue_lowres_image.png",
                     img_path2 = "data/MusMusculus/sag_post_1/spatial/tissue_lowres_image.png",
                     slice1 = "Anterior",
                     slice2 = "Posterior",
                     cell_types_interest = hip_layers,
                     return_legend = TRUE,
                     img_alpha = 1,
                     scatterpie_alpha = 1,
                     pie_scale = 0.2,
                     col_df = col_df) %>%   
  ggpubr::ggexport(plotlist = list(.), 
                   filename = sprintf("%s/hippocampus_spatial_scatterpie_allen_%s_%s.pdf",
                                    an_mouse, id_comp, id_nmf), 
                   width = 12, 
                   height = 9, 
                   res = 600)

```

##### Layer 2 neurons
```{r}
l2_neurons <- c("L2.3.IT.Cdc14a", "L2.3.IT.Cxcl14", "L2.3.IT.Ndst4.Endou", "L2.3.IT.Otof", "L2.3.IT.Plch1")

l2_scatterpie <- joint_scatterpie_fun(se_obj = brain, 
                     cell_types_all = cell_types_metadata,
                     img_path1 = "data/MusMusculus/sag_ant_1/spatial/tissue_lowres_image.png",
                     img_path2 = "data/MusMusculus/sag_post_1/spatial/tissue_lowres_image.png",
                     slice1 = "Anterior",
                     slice2 = "Posterior",
                     cell_types_interest = l2_neurons,
                     return_legend = FALSE,
                     img_alpha = 1,
                     scatterpie_alpha = 1,
                     pie_scale = 0.2,
                     col_df = col_df, 
                     arr_leg = FALSE) + 
                  ggtitle("All L2/3 neuron subtypes") + 
                  theme(
                    plot.margin = margin(1.5, 0, 1.5, 0, "cm"), 
                    plot.title = element_text(hjust = 0.5, size = 20, face = "bold")
                    )

cowplot::plot_grid(plotlist = list(l2_scatterpie))
ggpubr::ggexport(plotlist = list(l2_scatterpie), 
                 filename = sprintf("%s/layer2_spatial_scatterpie_allen_%s_%s.pdf",
                                  an_mouse, id_comp, id_nmf), 
                 width = 12, 
                 height = 9, 
                 res = 600)

```

###### Image composition
Here we are going to do an image composition of the plots with ISH data
```{r}
l2_df <- data.frame(colnm = l2_neurons, 
                    pltnm = c("L2/3 IT Cdc14a", "L2/3 IT Cxcl14", "L2/3 IT Ndst4 Endou", "L2/3 IT Otof", "L2/3 IT Plch1"))

# ct <- "L2.3.IT.Cxcl14"
l2_plots <- lapply(l2_neurons, function(ct) {
  print(ct)
  ct_scatterpie <- joint_scatterpie_fun(se_obj = brain, 
                     cell_types_all = cell_types_metadata,
                     img_path1 = "data/MusMusculus/sag_ant_1/spatial/tissue_lowres_image.png",
                     img_path2 = "data/MusMusculus/sag_post_1/spatial/tissue_lowres_image.png",
                     slice1 = "Anterior",
                     slice2 = "Posterior",
                     cell_types_interest = ct,
                     return_legend = TRUE,
                     img_alpha = 1,
                     scatterpie_alpha = 1,
                     pie_scale = 1,
                     col_df = col_df, 
                     arr_leg = FALSE) %>% 
    ggpubr::annotate_figure(.,
                top = ggpubr::text_grob(l2_df[l2_df$colnm == ct, "pltnm"], 
                                        face = "bold", 
                                        size = 14))

  ish_path_ls <- list.files(path = sprintf("%s/img", an_mouse), 
                       pattern = sprintf("%s.*ISH-crop_lowres.jpg",ct), 
                       full.names = TRUE)
  
  ish_plot_ls <- lapply(ish_path_ls, function(path) {
    plot_image(img_path = path)
  })
  
  ish_arr <- cowplot::plot_grid(plotlist = ish_plot_ls, 
                                align = "hv", 
                                nrow = 1, 
                                ncol = 2, 
                                axis = "trlb")

  exp_path_ls <- list.files(path = sprintf("%s/img", an_mouse), 
                       pattern = sprintf("%s.*Exp-crop.jpg",ct), 
                       full.names = TRUE)
  
  exp_plot_ls <- lapply(exp_path_ls, function(path) {
    plot_image(img_path = path)
  })
  exp_arr <- cowplot::plot_grid(plotlist = exp_plot_ls, 
                                align = "hv", 
                                nrow = 1, 
                                ncol = 2, 
                                axis = "trlb")
  
  tmp_arr <- cowplot::plot_grid(plotlist = list(ct_scatterpie, exp_arr, ish_arr),
                     nrow = 3, 
                     ncol = 1, 
                     align = "hv") + 
    theme(panel.border = element_rect(color = "black", fill = NA, size = 1))
  return(tmp_arr)
})


saveRDS(l2_plots, sprintf("%s/%s/l2_image_plots_%s_%s.RDS", an_mouse, robj_dir, id_comp, id_nmf))
l2_plots <- readRDS(sprintf("%s/%s/l2_image_plots_%s_%s.RDS", an_mouse, robj_dir, id_comp, id_nmf))

bottom_row <- cowplot::plot_grid(plotlist = list(l2_plots[[1]],
                                                 NULL, 
                                                 l2_plots[[2]],
                                                 NULL, 
                                                 l2_plots[[3]]),
                   ncol = 5,
                   nrow = 1,
                   rel_widths = c(1, 0.68, 1, 0.68, 1),
                   # align = "hv",
                   axis = "trbl", 
                   labels = c("D", "", "E", "", "F"))

top_row <- cowplot::plot_grid(plotlist = list(l2_plots[[4]],
                                              NULL,
                                              l2_scatterpie + theme(plot.background = element_rect(colour = "black", fill=NA, size=2)),
                                              NULL,
                                              l2_plots[[5]]),
                   ncol = 5,
                   nrow = 1,
                   rel_widths = c(0.5, 0.1, 1, 0.1, 0.5),
                   # align = "hv",
                   axis = "trbl", 
                   labels = c("A", "", "B", "", "C"))

l2_def <- cowplot::plot_grid(plotlist = list(NULL, NULL, NULL, 
                                   NULL, top_row, NULL, 
                                   NULL, NULL, NULL,
                                   NULL, bottom_row, NULL,
                                   NULL, NULL, NULL), 
                   ncol = 3, 
                   nrow = 5, 
                   rel_heights = c(0.05, 1, 0.1, 1, 0.05),
                   rel_widths = c(0.05, 1, 0.05))

mouse_plt_dir <- sprintf("%s/%s", an_mouse, plt_dir); 
dir.create(path = mouse_plt_dir, showWarnings = FALSE, recursive = TRUE)
ggpubr::ggexport(plotlist = list(l2_def),
                 filename = sprintf("%s/Figure_layer2_spatial_scatterpie_allen_%s_%s.pdf",
                                  mouse_plt_dir, id_comp, id_nmf), 
                 width = 18, 
                 height = 12)
```
