---
title: "Mouse brain supplementary figures"
author: "Marc Elosua-Bayes"
date: "4/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(purrr)
library(SPOTlight)
library(NMF)
library(nnls)
library(scrattch.io)
library(cowplot)
library(imager)
library(scatterpie)
library(tidyverse)
library(flextable)
library(webshot)
```

## Introduction

This is an Rmarkdown document where we are going to build the supplementary figures related to the brain.
We are going to plot:
* Table with number of cells per cell type
* Topic profile reproducibility
* Plot showing the location of all the cell types one by one

## Load data
Paths
```{r}
tech <- "sc"
tissue <- "allen_ref_70k"
dwn_smplng <- "both"
org <- "mm"
source("misc/paths_vrs.R")
```

Set common parameters
```{r}
clust_vr <- "subclass_label"
cl_n <- 100
method <- "nsNMF"
transf <- "uv"
hvg <- 3000
FC <- 1
pct1 <- 0.9

id_nmf <- sprintf("cln-%s_transf-%s_method-%s_hvg-%s_FC-%s_pct1-%s", 
                  cl_n, transf, method, hvg, FC, pct1)
data_dir <- "data/MusMusculus/allen_reference"
options(stringsAsFactors = FALSE)


library(RColorBrewer)
n <- 60
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector  <-  unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

## Load data
```{r}
brain <- readRDS(sprintf("%s/%s/brain1_processed.RDS", an_mouse, robj_dir))

allen_ref_70k <- readRDS(file = "data/MusMusculus/allen_reference/allen_ref_70k_processed.RDS")

decon_mtrx <- readRDS(file = sprintf("decon_mtrx_%s_%s.RDS", id_comp, id_nmf))
```

Join the data so its easier to work with
```{r}
cell_types <- colnames(decon_mtrx)[which(colnames(decon_mtrx) != "res_ss")]

brain@meta.data <- cbind(brain@meta.data, decon_mtrx)
cell_types_metadata <- colnames(brain@meta.data)[colnames(brain@meta.data) %in% cell_types]
```

To maintain consistent colors for each cell type between different plots we will create a dataframe of equivalencies which we will draw colors from
```{r}
col_df <- data.frame(cell_types = cell_types,
           col_vector = col_vector[1:length(cell_types)])
```

Dataframe with plot names
```{r}
colnames(decon_mtrx)
ct_names <- c("Astro", "CA1sp", "CA1sp.SUB-sp.Kcnip1", "CA2sp.IG", "CA3sp", 
              "Car3", "CR", "DG", "Doublet", "Endo", "IT.RHP.Dcn", 
              "L2.3.IT.Cdc14a", "L2.3.IT.Cxcl14", "L2.3.IT.Ndst4.Endou", 
              "L2.3.IT.Otof", "L2.3.IT.Plch1", "L4.IT", "L4.5.IT", "L5.ET", 
              "L5.IT", "L6.CT" ,"L6.IT", "L6b", "Lamp5", "Lamp5.Lhx6", 
              "Low.Quality", "Ly6g6e", "Macrophage", "Meis2", "NP", "Oligo",
              "PIR.Six3", "POST-PRE-PAR.Ptgfr", "Pvalb", "RHP.Cplx3", 
              "RSP_ACA.IT.Scnn1a", "RSP.ACA.L4.5.IT", "Serpinf1", "SMC", "Sncg",
              "Sncg.Ndnf.HPF", "Sst", "Sst.Chodl", "SUB-Sp.Ndst4", "Unknown", 
              "Vip", "VLMC")

plt_names <- c("Astrocytes", "CA1sp", "CA1sp/SUB-sp Kcnip1", "CA2sp/IG", "CA3sp",
               "Car3", "Cajal-Retzius", "DG", "Doublet", "Endothelium", 
               "IT RHP Dcn", "L2/3 IT Cdc14a", "L2/3 IT Cxcl14", 
               "L2/3 IT Ndst4 Endou", "L2/3 IT Otof", "L2/3 IT Plch1", "L4 IT", 
               "L4/5 IT", "L5 ET", "L5 IT", "L6 CT" ,"L6 IT", "L6b", "Lamp5", 
               "Lamp5 Lhx6", "Low quality", "Ly6g6e", "Macrophage", "Meis2", 
               "Neural Progenitor", "Oligodendrocyte", "PIR Six3", 
               "POST-PRE-PAR Ptgfr", "Pvalb", "RHP Cplx3", "RSP/ACA IT Scnn1a", 
               "RSP/ACA L4/5 IT", "Serpinf1", "Smooth Muscle Cells", "Sncg", 
               "Sncg/Ndnf HPF", "Sst", "Sst Chodl", "SUB-Sp Ndst4", "Unknown", 
               "Vip", "VLMC")

ct_df <- data.frame(ct_names, plt_names)
```

### Cell type Table
```{r}
df_count <- data.frame(allen_ref_70k@meta.data) %>% 
  count(subclass_label)

colnames(df_count) <- make.unique()

count_wide <- cbind(df_count[1:16, ], df_count[17:32, ], df_count[33:48, ])
colnames(count_wide) <- make.unique(colnames(count_wide))
count_wide[is.na(count_wide)] <- ""

cell_type_table <- flextable(count_wide) %>%
  bold(j = c(1, 3, 5), 
       bold = TRUE, 
       part = "body") %>%
  bold(part = "header", 
       bold = TRUE ) %>% 
  set_header_labels(subclass_label = "Cell type",
                    n = "Number of cells",
                    subclass_label.1 = "Cell type",
                    n.1 = "Number of cells",
                    subclass_label.2 = "Cell type",
                    n.2 = "Number of cells",
                    subclass_label.3 = "Cell type",
                    n.3 = "Number of cells") %>%
  # align_nottext_col(align = "center") %>%
  fontsize(i = 1, 
           j = NULL, 
           size = 15, 
           part = "header") %>% 
  fontsize(i = NULL, 
           j = NULL, 
           size = 12, 
           part = "body") %>% 
  autofit() %>% 
  add_header_lines("Supplementary Table 1: Composition of the Allen Mouse Brain Atlas dataset.") %>% 
  # add_footer_lines("sp: stratum pyramidale; IG") %>%
  fontsize(i = NULL, 
           j = NULL, 
           size = 12, 
           part = "header") %>% 
  flextable::theme_zebra() %>%
  align(j = c(2, 4, 6), 
        part = "body", 
        align = "center")
  # theme_vanilla()
cell_type_table

save_as_image(cell_type_table, 
              path = sprintf("%s/%s/Supplementary_Table_1_Allen_Mouse_Brain_Atlas.pdf",
                             an_mouse, plt_dir),
              zoom = 1, 
              expand = 100)
```

### All cell types location

```{r}
join_spatial_plots <- function(spatial_obj, ct) {
  ###########################################################################################################
  #### This functions takes in a feature to plot in 2 spatial slides and returns it with a unique legend ####
  ###########################################################################################################
  library(ggplot2)
  library(ggpubr)
  library(cowplot)
  library(Seurat)
  
  # Set common legend theme
  legend_theme <-  theme(legend.text = element_text(colour = "#3b3a39", size = 10),
                       legend.title = element_text(colour = "#3b3a39", vjust = 1))
  
  orig_pt <- SpatialFeaturePlot(spatial_obj,
                   features = ct,
                   pt.size.factor = 1,
                   alpha = c(0, 1)) + 
    theme(plot.title = element_text(hjust = 0.5))
  
  ### Modify individual plots ###
  orig_pt_ant <- orig_pt[[1]] +
    cowplot::theme_map() + 
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))+ 
    scale_fill_gradientn(
      # colours = colorRamps::matlab.like2(10),
      colours = heat.colors(10, rev = TRUE),
      limits = c(0, 1)) +
    labs(title = "", 
       fill = "Proportion")
  
  orig_pt_post <- orig_pt[[2]] +
    labs(title = "", fill = ct) +
    cowplot::theme_map() + 
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) + 
    scale_fill_gradientn(
      # colours = colorRamps::matlab.like2(10),
      colours = heat.colors(10, rev = TRUE),
      limits = c(0, 1))

  return(list(orig_pt_ant, orig_pt_post))
}
```

```{r}
ct_arrange <- lapply(cell_types, function(ct) {
  
  ct_ls <- join_spatial_plots(spatial_obj = brain, ct = ct)
  plt_tmp <- cowplot::plot_grid(plotlist = list(ct_ls[[1]] + theme(legend.position = "none"), 
                                                ct_ls[[2]] + theme(legend.position = "none")), 
                                nrow = 1, 
                                ncol = 2, 
                                align = "hv",
                                labels = ct_df[ct_df$ct_names == ct, "plt_names"],
                                label_size = 15
                                ) + 
    # theme(plot.margin = unit(c(1,1,1,1), "line"))
    theme(plot.margin = unit(c(0.2,0.2,0.2,0.2), "line"))
  
  leg_grobtable <- get_legend(ct_ls[[1]])
  plt_titleless <- cowplot::plot_grid(plotlist = list(plt_tmp, leg_grobtable),
                     ncol = 2,
                     nrow = 1, rel_widths = c(1, 0.1))
  return(plt_tmp)
}) %>% 
  ggarrange(plotlist = ., ncol = 6, nrow = 9, align = "hv")

# ggsave(plot = patchwork::wrap_plots(ct_arrange), 
#        device = cairo_pdf,
#        file = "a4_output.pdf", 
#        # width = 210, 
#        # height = 297, 
#        width = 420,
#        height = 600,
#        units = "mm")

cowplot::save_plot(plot =  ct_arrange,
                   filename = sprintf("%s/%s/Supplementary_Figure_AAA_all_cell_types_grid_%s_%s.svg",
                                      an_mouse, plt_dir, id_comp, id_nmf), 
                   base_width = 24,
                   base_height = 24)

ggpubr::ggexport(plotlist = list(ct_arrange),
                 filename = sprintf("%s/%s/Supplementary_Figure_AAA_all_cell_types_grid_%s_%s.jpeg",
                               an_mouse, plt_dir, id_comp, id_nmf),
                 width = 5000,
                 height = 4500,
                 res = 300)

```


### Topic profile reproducibility
Load cluster markers
```{r}
cluster_markers_all <- readRDS(file = "data/MusMusculus/allen_reference/markers_allen_ref_70k.RDS")

cluster_markers_filt <- cluster_markers_all %>% 
  filter(avg_logFC > FC & pct.1 > pct1)
```

Downsample + train model
```{r}
id_nmf_iter <- sprintf("cln-30_transf-%s_method-%s_hvg-%s_FC-%s_pct1-%s",
                     transf, method, hvg, FC, pct1)

set.seed(123)
nmf_mod_ls_10 <- lapply(seq_len(10), function(i) {
  print(i)
  print("Downsmapling")
  allen_ref_70k_down <- downsample_se_obj(se_obj = allen_ref_70k,
                                          clust_vr = clust_vr,
                                          cluster_markers = cluster_markers_filt, 
                                          cl_n = 30, 
                                          hvg = hvg)
  
  print("Training model")
  nmf_mod_ls <- train_nmf(cluster_markers = cluster_markers_filt, 
                          se_sc = allen_ref_70k_down, 
                          mtrx_spatial = brain@assays$Spatial@counts, 
                          ntop = NULL, 
                          transf = transf, 
                          clust_vr = clust_vr, 
                          method = "nsNMF")

  saveRDS(object = nmf_mod_ls, 
        file = sprintf("%s/%s/nmf_mod_ls_10_%s_%s_iter-%s.rds", 
                       an_mouse, robj_dir, id_comp, id_nmf_iter, i))

  return(nmf_mod_ls)
})


```

Topic profile plots
```{r}
topic_plt_arr <- lapply(seq_len(10), function(i) {
  nmf_mod_ls <- readRDS(file = sprintf("%s/%s/nmf_mod_ls_10_%s_%s_iter-%s.rds", 
                     an_mouse, robj_dir, id_comp, id_nmf_iter, i))

  nmf_mod <- nmf_mod_ls[[1]]

  # get matrix H
  h <- coef(nmf_mod)
  rownames(h) <- paste("Topic", 1:nrow(h), sep = " ")
  
  topic_profile_plts <- dot_plot_profiles_fun(h = h,
                      train_cell_clust = nmf_mod_ls[[2]])
  
  tmp_plt <- topic_profile_plts[[2]] +
    scale_size(limits = c(0, 1)) +
    labs(title = sprintf("Topic Profiles: Iteration %s", i)) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 15),
      axis.text.x = element_text(angle = 90, hjust = 1, size = 10),
      axis.text.y = element_text(size = 10))
  
  return(tmp_plt)
}) %>% ggpubr::ggarrange(plotlist = ., ncol = 3, nrow = 4)


cowplot::save_plot(
  filename = sprintf("%s/%s/Supplementary_Figure_FFF_topic_signatures_arrange_10_%s_%s_.pdf", 
                     an_mouse, plt_dir, id_comp, id_nmf_iter),
  plot = topic_plt_arr,
  base_width = 24,
  base_height = 30)

cowplot::save_plot(
  filename = sprintf("%s/%s/Supplementary_Figure_FFF_topic_signatures_arrange_10_%s_%s_.svg",
                     an_mouse, plt_dir, id_comp, id_nmf_iter),
  plot = topic_plt_arr,
   base_width = 24,
   base_height = 30)


cowplot::save_plot(
  filename = "~/Desktop/SPOTlight_plots/topic_profiles.png",
  plot = topic_plt_arr,
  base_width = 36,
  base_height = 30)

```

Show topic profiles of all the cell types within a cell type
```{r}
i <- 1
nmf_mod_ls <- readRDS(file = sprintf("%s/%s/nmf_mod_ls_10_%s_%s_iter-%s.rds", 
                   an_mouse, robj_dir, id_comp, id_nmf_iter, i))

nmf_mod <- nmf_mod_ls[[1]]
```

Extract matrices form the model:
```{r}
# get matrix W
w <- basis(nmf_mod)
dim(w)

# get matrix H
h <- coef(nmf_mod)
dim(h)
rownames(h) <- paste("Topic", 1:nrow(h), sep = " ")

```

Save all cell topic profiles
```{r}
topic_profile_plts <- dot_plot_profiles_fun(h = h,
                      train_cell_clust = nmf_mod_ls[[2]],
                      clust_vr = clust_vr)
all_cell_prof_plt <- topic_profile_plts[[1]] + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 2)) +
  scale_size(range = c(0, 3))

ggpubr::ggexport(plotlist = list(all_cell_prof_plt),
                 filename = sprintf("%s/%s/Supplementary_Figure_GGG_all_cell_topic_profile_%s_%s.pdf", 
                                    an_mouse, plt_dir, id_comp, id_nmf_iter),
                 width = 32,
                 height = 32,
                 res = 600)

cowplot::save_plot(plot = all_cell_prof_plt,
                 filename = sprintf("%s/%s/Supplementary_Figure_GGG_all_cell_topic_profile_%s_%s.svg", 
                                    an_mouse, plt_dir, id_comp, id_nmf_iter),
                 base_width = 32,
                 base_height = 32)


ggpubr::ggexport(plotlist = list(all_cell_prof_plt),
                 filename = sprintf("%s/%s/Supplementary_Figure_GGG_all_cell_topic_profile_%s_%s.jpeg",
                                    an_mouse, plt_dir, id_comp, id_nmf_iter),
                 width = 5000,
                 height = 5000,
                 res = 400)

```
