---
title: "Synthetic spot generator"
author: "Marc Elosua-Bayes"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
In this R markdown document we will generate a common set of 1.000 common synthetic mixtures on which we will test the different deconvolution tools. In order to do this we will use the PBMC dataset used for the benchmarking. We will use the data from the QUARTZ-seq dataset since it show to be the highest quality.

## Libraries
```{r}
library(SPOTlight)
```

Parameters
```{r}
set.seed(321)

tech <- "QUARTZ-seq"
org <- "hs"
tissue <- "pbmc"
dwn_smplng <- "both"
source(here::here("misc/paths_vrs.R"))

if (! file.exists(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)))) {
  dir.create(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
  dir.create(here::here(sprintf("analysis/tool_benchmarking/%s", robj_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
}
```

## Load data
```{r}
load(here::here("data/techs/sce.all_classified.technologies.RData"))
# Loads object named sce
```

## Subset QUARTZ-seq
Select data from QUARTZ-seq only to generate the synthetic mixtures.
```{r}
sce_quartz <- sce[, sce$batch == "Quartz-Seq2"]
sce_quartz$nnet2 <- as.character(sce_quartz$nnet2)
table(sce_quartz$batch)
table(sce_quartz$nnet2)
```

Since there just 1 Megakaryocytes we are going to remove it from the dataset
```{r}
sce_quartz <- sce_quartz[, sce_quartz$nnet2 != "Megakaryocytes"]
```

Convert SCE to Seurat object
```{r}
se_quartz <- Seurat::CreateSeuratObject(counts = sce_quartz@assays$data$counts,
                                        meta.data = data.frame(sce_quartz@colData))

saveRDS(object = se_quartz,
        file = here::here(sprintf("%s/%s/se_quartz.RDS", an_tools, robj_dir)))
```

## Generate synthetic mixtures
To this end we use `SPOTlight::test_spot_fun` which allows us to generate mixtures resembling visium's spots with average transcriptome sizes of 20K UMI. If the mixture of cells randomly generated exceeds 25K UMI we randomly downsample to 20k.
```{r}
synthetic_mixtures <- SPOTlight::test_spot_fun(se_obj = se_quartz,
                                               clust_vr = "nnet2",
                                               n = 1000,
                                               verbose = TRUE)

saveRDS(object = synthetic_mixtures,
        file = here::here(sprintf("%s/%s/common_synthetic_mixtures.RDS", an_tools, robj_dir)))
```

## Generate CIBERSORTx reference set
```{r}
#### Get cell IDs to subset by cluster ####
keep_ids <- lapply(split(se_quartz@meta.data, se_quartz@meta.data[, "nnet2"]), function(subdf) {
  # Determine n_sample, if the size of the group is < cl_n use all the group, if not just use cl_n
  n_sample <- if_else(nrow(subdf) < 5, as.numeric(nrow(subdf)), as.numeric(5))
  # Subset a random selection of that group and get the identifiers
  tmp_ds <- subdf[sample(seq_len(nrow(subdf)), n_sample), ] %>%
    tibble::rownames_to_column("ID") %>%
    dplyr::pull(ID)
  return(tmp_ds)
}) %>%
  purrr::flatten_chr() # flatten the list into a vector

#### Subset seurat object ####
se_obj <- se_quartz[, keep_ids]
mtrx <- as.matrix(se_obj@assays$RNA@counts)
colnames(mtrx) <- se_obj$nnet2
readr::write_tsv(x = data.frame(mtrx) %>% tibble::rownames_to_column("Gene"),
                 path = here::here(sprintf("%s/%s/cibersortx_signature.tsv", an_tools, robj_dir)))
```

## Select first 50 mixtures
```{r}
synthetic_mixtures <- readRDS(file = here::here(sprintf("%s/%s/common_synthetic_mixtures.RDS", an_tools, robj_dir)))
readr::write_tsv(x = data.frame(synthetic_mixtures[[1]][, 1:50]) %>% tibble::rownames_to_column("Gene"),
                 path = here::here(sprintf("%s/%s/cibersortx_buks.tsv", an_tools, robj_dir)))
```

