---
title: "DeconRNAseq deconvolution"
author: "Marc Elosua-Bayes"
date: "7/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
In this Rmarkdown document we are going to use DeconRNAseq to deconvolute mixtures of cells. Ultimately the goal is to compare the performance with SPOTlight and other deconvolution tools. We use a consesus set of 1000 spots made out of between 2-8 cells reflecting the composition of Visium's spot. Synthetic mixtures and scRNAseq datasets are common for all the tools and are generated in the script common_synthetic_spots.Rmd

## Libraries
Install SCDC package
https://github.com/meichendong/SCDC
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

BiocManager::install("DeconRNASeq")
library(DeconRNASeq)
```

```{r}
library(SPOTlight)
library(Seurat)
library(dplyr)
library(ggplot2)
library(Matrix)
```

## Parameters
```{r}
set.seed(321)
tech <- "QUARTZ-seq"
org <- "hs"
tissue <- "pbmc"
dwn_smplng <- "both"
source(here::here("misc/paths_vrs.R"))

if (! file.exists(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)))) {
  dir.create(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
  dir.create(here::here(sprintf("analysis/tool_benchmarking/%s", robj_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
}
```

## Load data
Here we are loading the scRNAseq seurat object from QUARTZ-seq and the common synthetic spots. Both generated in the common_synthetic_spots.Rmd markdown document in this folder.
```{r}
se_quartz <- readRDS(file = here::here(sprintf("%s/%s/se_quartz.RDS", an_tools, robj_dir)))

synthetic_mixtures <- readRDS(file = here::here(sprintf("%s/%s/common_synthetic_mixtures.RDS", an_tools, robj_dir)))
```

## DeconRNAseq deconvolution
Create signatures averaging the transcriptomes of 50 random cells per cell types. If one cell type doesn't have enough cell it grabs all of them.
```{r}
ct <- "HEK cells"
sign_mtrx <- lapply(unique(se_quartz$nnet2), function(ct) {
  
  if (sum(se_quartz$nnet2 == ct) > 50) n <- 50 else n <- sum(se_quartz$nnet2 == ct)
  
  keep_id <- base::sample(x = colnames(se_quartz[, se_quartz$nnet2 == ct]), size = n)
  tmp_se <- se_quartz[, keep_id]
  
  mtrx <- as.matrix(rowMeans(tmp_se@assays$RNA@counts))
  colnames(mtrx) <- ct
  return(mtrx)
}) %>% dplyr::bind_cols() %>% as.data.frame()

colnames(sign_mtrx) <- unique(se_quartz$nnet2)
```

carry out deconvolution
```{r}
deconrnaseq_ls <- DeconRNASeq::DeconRNASeq(datasets = as.data.frame(synthetic_mixtures[[1]]),
                         signatures = sign_mtrx)
saveRDS(object = deconrnaseq_ls,
        file = here::here(sprintf("%s/%s/deconrnaseq_deconv.RDS", an_tools, robj_dir)))
```

## Assess performance
```{r}
synthetic_comp <- as.matrix(synthetic_mixtures[[2]] / rowSums(synthetic_mixtures[[2]]))

colnames(deconrnaseq_ls[["out.all"]]); colnames(synthetic_comp)
SPOTlight::test_synthetic_performance(test_spots_metadata_mtrx = deconrnaseq_ls[["out.all"]],
                                      spot_composition_mtrx = synthetic_comp)

```

