---
title: "MuSiC deconvolution"
author: "Marc Elosua-Bayes"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Introduction
In this Rmarkdown document we are going to use MuSiC to deconvolute mixtures of cells. Ultimately the goal is to compare the performance with SPOTlight and other deconvolution tools. We use a consesus set of 1000 spots made out of between 2-8 cells reflecting the composition of Visium's spot. Synthetic mixtures and scRNAseq datasets are common for all the tools and are generated in the script common_synthetic_spots.Rmd

## Libraries
Install MuSiC from GitHub
https://xuranw.github.io/MuSiC/articles/MuSiC.html
```{r}
# install the MuSiC package
devtools::install_github('xuranw/MuSiC')

# load
library(MuSiC)
```

```{r}
library(SPOTlight)
library(Seurat)
library(dplyr)
library(ggplot2)
library(Matrix)
```

## Parameters
```{r}
set.seed(321)
tech <- "QUARTZ-seq"
org <- "hs"
tissue <- "pbmc"
dwn_smplng <- "both"
source(here::here("misc/paths_vrs.R"))

if (! file.exists(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)))) {
  dir.create(sprintf("analysis/%s/%s", an_tools, plt_dir),
             showWarnings = FALSE,
             recursive = TRUE)
  dir.create(here::here(sprintf("analysis/tool_benchmarking/%s", robj_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
}
```

## Load data
Here we are loading the scRNAseq seurat object from QUARTZ-seq and the common synthetic spots. Both generated in the common_synthetic_spots.Rmd markdown document in this folder.
```{r}
se_quartz <- readRDS(file = here::here(sprintf("%s/%s/se_quartz.RDS", an_tools, robj_dir)))

synthetic_mixtures <- readRDS(file = here::here(sprintf("%s/%s/common_synthetic_mixtures.RDS", an_tools, robj_dir)))
```


## Run MuSiC
1st we need to conver our data to ExpressionSet format
```{r}
se_quartz[["barcode"]] <- colnames(se_quartz)
expr_sc <- Biobase::ExpressionSet(assayData = as.matrix(se_quartz@assays$RNA@counts),
                                  phenoData = Biobase::AnnotatedDataFrame(data = se_quartz@meta.data))

expr_mix <- Biobase::ExpressionSet(assayData = as.matrix(synthetic_mixtures[[1]]))
```

Run deconvolution
```{r}
strt <- Sys.time()
library(xbioc)
music_deconv <- MuSiC::music_prop(bulk.eset = expr_mix,
                                  sc.eset = expr_sc,
                                  markers = NULL,
                                  clusters = "nnet2",
                                  samples = "barcode",
                                  select.ct = NULL,
                                  verbose = FALSE)
tot_time <- difftime(time1 = Sys.time(),
                     time2 = strt,
                     units = "min")

print(glue::glue("MuSiC took {round(tot_time, 2)} minutes to carry out the deconvolution"))

saveRDS(object = music_deconv,
        file = here::here(sprintf("%s/%s/music_deconv.RDS", an_tools, robj_dir)))
```

Extract parts and assess performance
```{r}
music_pred <- music_deconv[["Est.prop.weighted"]]
synthetic_comp <- synthetic_mixtures[[2]] / rowSums(synthetic_mixtures[[2]])

SPOTlight::test_synthetic_performance(test_spots_metadata_mtrx = music_pred,
                                      spot_composition_mtrx = as.matrix(synthetic_comp))

is(music_pred)
is(synthetic_comp)
```


