---
title: "RCTD deconvolution"
author: "Marc Elosua-Bayes"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
In this Rmarkdown document we are going to use RCTD to deconvolute mixtures of cells. Ultimately the goal is to compare the performance with SPOTlight and other deconvolution tools. We use a consesus set of 1000 spots made out of between 2-8 cells reflecting the composition of Visium's spot. Synthetic mixtures and scRNAseq datasets are common for all the tools and are generated in the script common_synthetic_spots.Rmd

RCTD is developed by Dylan M. Cable from Rafa Irizarry's lab, the preprint can be found [here](https://www.biorxiv.org/content/10.1101/2020.05.07.082750v1). 

## Libraries
Install RCTD package
https://github.com/dmcable/RCTD
```{r}
# install.packages("devtools")
# devtools::install_github("dmcable/RCTD", build_vignettes = TRUE)
library(RCTD)
library(Seurat)
library(dplyr)
```

## Parameters
```{r}
set.seed(321)

tech <- "QUARTZ-seq"
org <- "hs"
tissue <- "pbmc"
dwn_smplng <- "both"
source(here::here("misc/paths_vrs.R"))

if (! file.exists(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)))) {
  dir.create(sprintf("analysis/%s/%s", an_tools, plt_dir),
             showWarnings = FALSE,
             recursive = TRUE)
  dir.create(here::here(sprintf("analysis/tool_benchmarking/%s", robj_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
}
```

## Load data
Here we are loading the scRNAseq seurat object from QUARTZ-seq and the common synthetic spots. Both generated in the common_synthetic_spots.Rmd markdown document in this folder.
```{r}
se_quartz <- readRDS(file = here::here(sprintf("%s/%s/se_quartz.RDS", an_tools, robj_dir)))

synthetic_mixtures <- readRDS(file = here::here(sprintf("%s/%s/common_synthetic_mixtures.RDS", an_tools, robj_dir)))
```

## Prep data
vignette [here](https://raw.githack.com/dmcable/RCTD/dev/vignettes/spatial-transcriptomics.html)

### scRNAseq reference
```{r}
RCTD_structure <- function(sc_obj, clust_vr) {
  
  sc_obj[["Name"]] = sc_obj@meta.data[, clust_vr]
  
  # Cell type dictionary between cluster and cell-type
  ct <- unique(sc_obj@meta.data[, clust_vr])
  df_ct <- data.frame("Cluster" = 1:length(ct),
                      "Name" = ct)
  
  # 
  metadata <- sc_obj@meta.data %>%
    # Rownames to columns must be before left join since after it the rownames are erased
    tibble::rownames_to_column("barcode") %>%
    dplyr::left_join(df_ct, by = c("Name" = "Name")) %>%
    # Change names to “barcode”, “cluster”, “nUMI”
    mutate(
      cluster = Cluster,
      nUMI = nCount_RNA
      ) %>%
    dplyr::select(barcode, cluster, nUMI)
  
  expr_mtrx <- sc_obj@assays$RNA@counts
  
  return(list("meta_data" = metadata,
              "cell_type_dict" = df_ct,
              "dge" = expr_mtrx))
}
```

```{r eval = FALSE}
dir.create(here::here("analysis/tool_benchmarking/RCTD_data/reference"))
sc_ls <- RCTD_structure(sc_obj = se_quartz,
               clust_vr = "nnet2")

readr::write_csv(x = sc_ls[[1]],
                 path = here::here("analysis/tool_benchmarking/RCTD_data/reference/meta_data.csv"))

readr::write_csv(x = sc_ls[[2]],
                 path = here::here("analysis/tool_benchmarking/RCTD_data/reference/cell_type_dict.csv"))

sc_ls[[3]] %>%
  data.frame() %>%
  tibble::rownames_to_column("gene") %>%
  readr::write_csv(x = .,
                   path = here::here("analysis/tool_benchmarking/RCTD_data/reference/dge.csv"),
                   col_names = TRUE)

```

*Of note, here go to analysis/tool_benchmarking/RCTD_data/reference/dge.csv and remove the column name gene by "". If you don't it will detect the column as a gene and RCTD::dgeToSeurat won't be able to load the data*.

### Spatial data
```{r eval = FALSE}
dir.create(here::here("analysis/tool_benchmarking/RCTD_data/spatial"))
tmp <- synthetic_mixtures[[1]] %>%
  data.frame() %>%
  tibble::rownames_to_column("gene") %>% 
  dplyr::mutate(spot_0 = spot_1) %>%
  dplyr::select(gene, spot_0, everything()) %>% 
  readr::write_csv(x = ,
                   file = here::here("analysis/tool_benchmarking/RCTD_data/spatial/MappedDGEForR.csv"),
                   col_names = TRUE)

# Create aritificial coordinates for the synthetic spots
nrow(synthetic_mixtures[[2]])

# Since we have 1000 spots we can create an array of 20 * 50 matrix
coord <- expand.grid(1:20, 1:50)
colnames(coord) <- c("xcoord", "ycoord")
df_coord <- data.frame("barcodes" = paste("spot",1:1000, sep = "_"), coord)

readr::write_csv(x = df_coord,
                 file = here::here("analysis/tool_benchmarking/RCTD_data/spatial/BeadLocationsForR.csv"))
```

## RCTD deconvolution
### Read data in RCTD
Read data in RCTD format
```{r}
reference <- RCTD::dgeToSeurat(refdir = here::here("analysis/tool_benchmarking/RCTD_data/reference"))
puck <- RCTD::read.SpatialRNA(datadir = here::here("analysis/tool_benchmarking/RCTD_data/spatial"))
```

### Creating and running RCTD
We are now ready to create an RCTD object using the create.RCTD function. We simply need to pass in the SpatialRNA and scRNA-seq objects. There are several configuration options that can be set with this function:
```{r}
myRCTD <- RCTD::create.RCTD(spatialRNA = puck,
                            reference = reference,
                            max_cores = 1,
                            CELL_MIN = 18)
```

Now, we are ready to run RCTD, using the run.RCTD function. This function is equivalent to sequentially running the functions fitBulk, choose_sigma_c, and fitPixels. The doublet_mode argument sets whether RCTD will be run in ‘doublet mode’ (at most 1-2 cell types per pixel) or ‘full mode’ (no restrictions on number of cell types).
```{r}
myRCTD <- RCTD::run.RCTD(RCTD = myRCTD,
                         doublet_mode = FALSE)
```

### Assess results
The results of RCTD are located in the @results field. Of particular interest is @results$weights, a data frame of cell type weights for each pixel (for full mode).
```{r}
results <- myRCTD@results
```

normalize the cell type proportions to sum to 1.
```{r}
norm_weights <- sweep(results$weights, 1, rowSums(results$weights), '/') 
rctd_deconv <- myRCTD@results$weights
cell_type_names <- myRCTD@cell_type_info$info[[2]] #list of cell type names
```

### Save results
```{r}
saveRDS(object = myRCTD,
        file = sprintf("%s/%s/rctd_deconv.RDS", an_tools, robj_dir))
```

