---
title: "SPOTlight deconvolution"
author: "Marc Elosua-Bayes"
date: "7/15/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction
In this Rmarkdown document we are going to use SPOTlight to deconvolute mixtures of cells. Ultimately the goal is to compare the performance with SPOTlight and other deconvolution tools. We use a consesus set of 1000 spots made out of between 2-8 cells reflecting the composition of Visium's spot. Synthetic mixtures and scRNAseq datasets are common for all the tools and are generated in the script common_synthetic_spots.Rmd

## Libraries
```{r}
library(SPOTlight)
library(dplyr)
library(ggplot2)
library(Matrix)
```

## Parameters
```{r}
set.seed(321)
tech <- "QUARTZ-seq"
org <- "hs"
tissue <- "pbmc"
dwn_smplng <- "both"
source(here::here("misc/paths_vrs.R"))

if (! file.exists(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)))) {
  dir.create(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
  dir.create(here::here(sprintf("analysis/tool_benchmarking/%s", robj_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
}
```

## Load data
Here we are loading the scRNAseq seurat object from QUARTZ-seq and the common synthetic spots. Both generated in the common_synthetic_spots.Rmd markdown document in this folder.
```{r}
se_quartz <- readRDS(file = here::here(sprintf("%s/%s/se_quartz.RDS", an_tools, robj_dir)))

synthetic_mixtures <- readRDS(file = here::here(sprintf("%s/%s/common_synthetic_mixtures.RDS", an_tools, robj_dir)))
```

## Run SPOTlight
1st we need to find the marker genes for each cell type
```{r}
se_quartz <- Seurat::SCTransform(object = se_quartz,
                                 assay = "RNA",
                                 verbose = FALSE)

Seurat::Idents(se_quartz) <- se_quartz$nnet2
marker_genes <- Seurat::FindAllMarkers(object = se_quartz,
                                       assay = "SCT",
                                       slot = "data",
                                       min.pct = 0,
                                       only.pos = TRUE,
                                       logfc.threshold = 0)

marker_genes %>% dplyr::count(cluster)

marker_genes_filt <- marker_genes %>%
  dplyr::filter(pct.1 > 0.9)

marker_genes_filt %>% dplyr::count(cluster)
```

Run SPOTlight
```{r eval = FALSE}
spotlight_ls <- SPOTlight::spotlight_deconvolution(se_sc = se_quartz,
                                   counts_spatial = synthetic_mixtures[[1]],
                                   clust_vr = "nnet2",
                                   cluster_markers = marker_genes,
                                   cl_n = 100,
                                   hvg = 3000,
                                   ntop = NULL,
                                   transf = "uv",
                                   method = "nsNMF",
                                   min_cont = 0.09,
                                   assay = "RNA",
                                   slot = "counts")

saveRDS(object = spotlight_ls,
        file = here::here(sprintf("%s/%s/spotlight_deconv.RDS", an_tools, robj_dir)))
```


```{r fig.height=8, fig.width=8}
spotlight_ls <- readRDS(file = here::here(sprintf("%s/%s/spotlight_deconv.RDS",
                                                  an_tools, robj_dir)))

h <- NMF::coef(spotlight_ls[[1]][[1]])
train_labs <- spotlight_ls[[1]][[2]]

profile_plt <- SPOTlight::dot_plot_profiles_fun(
  h = h,
  train_cell_clust = train_labs)

profile_plt[[2]]
```

Extract parts and assess performance
```{r}
spotlight_deconv <- spotlight_ls[[2]][, colnames(spotlight_ls[[2]]) != "res_ss"]
synthetic_comp <- as.matrix(synthetic_mixtures[[2]] /
                              rowSums(synthetic_mixtures[[2]]))

colnames(spotlight_deconv); colnames(synthetic_comp)
SPOTlight::test_synthetic_performance(
  test_spots_metadata_mtrx = spotlight_deconv[, colnames(synthetic_comp)],
  spot_composition_mtrx = synthetic_comp)
```

## Session Info
```{r}
sessionInfo()
```
