---
title: "Tool benchmark"
author: "Marc Elosua-Bayes"
date: "7/17/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
In this Rmarkdown document we will compare the performance of the different tools used to deconvolute mixtures.

## Libraries
```{r}
library(dplyr)
library(ggplot2)
library(RColorBrewer)
```

## Parameters
```{r}
tech <- "QUARTZ-seq"
org <- "hs"
tissue <- "pbmc"
dwn_smplng <- "both"
source(here::here("misc/paths_vrs.R"))

if (! file.exists(sprintf("analysis/%s/%s", an_tools, plt_dir))) {
  dir.create(here::here(sprintf("analysis/%s/%s", an_tools, plt_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
  dir.create(here::here(sprintf("analysis/tool_benchmarking/%s", robj_dir)),
             showWarnings = FALSE,
             recursive = TRUE)
}
```

Color palette to use for each technology
```{r}
techs <- c("SPOTlight", "MuSiC weighted", "MuSiC all genes", "SCDC", "NMFreg",
           "DeconRNAseq", "RCTD", "CIBERSORT", "CIBERSORTx", "SPOTlight-CoGAPS") # "deepSCore"

set1 <- RColorBrewer::brewer.pal(length(techs), "Set3")

col_df <- data.frame("Tool" = techs,
                     "color" = set1)
```

## Load data
Here we are loading the scRNAseq seurat object from QUARTZ-seq and the common synthetic spots. Both generated in the common_synthetic_spots.Rmd markdown document in this folder.
```{r}
se_quartz <- readRDS(file = here::here(sprintf("%s/%s/se_quartz.RDS", an_tools, robj_dir)))

synthetic_mixtures <- readRDS(file = here::here(sprintf("%s/%s/common_synthetic_mixtures.RDS", an_tools, robj_dir)))
synthetic_comp <- as.matrix(synthetic_mixtures[[2]] / rowSums(synthetic_mixtures[[2]]))
```

We also load the deconvolution outputs from the different tools.
```{r}
# MuSiC
music_deconv <- readRDS(file = here::here(sprintf("%s/%s/music_deconv.RDS",
                                                  an_tools, robj_dir)))
music_mtrx_w <- music_deconv[["Est.prop.weighted"]]
colnames(music_mtrx_w) <- gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
                                            x = colnames(music_mtrx_w),
                                            perl = TRUE)

music_mtrx_a <- music_deconv[["Est.prop.allgene"]]
colnames(music_mtrx_a) <- gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
                                            x = colnames(music_mtrx_a),
                                            perl = TRUE)
# SPOTlight
spotlight_deconv <- readRDS(
  file = here::here(sprintf("%s/%s/spotlight_deconv.RDS",
                            an_tools, robj_dir)))
spotlight_mtrx <- spotlight_deconv[[2]]

# SCDC
scdc_obj <- readRDS(
  file = here::here(sprintf("%s/%s/scdc_deconv.RDS",
                            an_tools, robj_dir)))

scdc_deconv <- scdc_obj$prop.est.mvw
colnames(scdc_deconv) <- gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
                                            x = colnames(scdc_deconv),
                                            perl = TRUE)

# NMFreg
nmfreg_deconv <- readRDS(
  file = here::here(sprintf("%s/%s/nmfreg_ls_deconv.RDS",
                            an_tools, robj_dir)))
nmfreg_mtrx <- nmfreg_deconv[[2]]

# deconRNAseq
deconrnaseq_deconv <- readRDS(
  file = here::here(sprintf("%s/%s/deconrnaseq_deconv.RDS",
                 an_tools, robj_dir)))
deconrnaseq_mtrx <- deconrnaseq_deconv[["out.all"]]
colnames(deconrnaseq_mtrx) <- gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
                                            x = colnames(deconrnaseq_mtrx),
                                            perl = TRUE)

# RCTD
myRCTD <- readRDS(
  file = here::here(sprintf("%s/%s/rctd_deconv.RDS",
                            an_tools, robj_dir)))
rctd_deconv <- myRCTD@results$weights

colnames(rctd_deconv) <- gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
                              x = colnames(rctd_deconv),
                              perl = TRUE)

# # CIBERSORT
# CIBERSORT_deconv <- readRDS(
#   file = here::here(sprintf("%s/%s/cibersort_deconv.RDS",
#                             an_tools, robj_dir)))

# CIBERSORTx
CIBERSORTx_deconv <- readr::read_tsv(
  file = here::here(sprintf("%s/%s/CIBERSORTx_Results.txt",
                            an_tools, robj_dir)),
  col_names = TRUE) %>%
  tibble::column_to_rownames("Mixture") %>%
  as.matrix()

colnames(CIBERSORTx_deconv) <- gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
                              x = colnames(CIBERSORTx_deconv),
                              perl = TRUE)

# deepSCore
# deepSCore_deconv <- readRDS(file = here::here(sprintf("%s/%s/deepSCore_deconv.RDS",
#                                   an_tools, robj_dir)))
# 
# colnames(deepSCore_deconv) <- gsub(pattern = "[[:punct:]]|[[:blank:]]", ".",
#                               x = colnames(deepSCore_deconv),
#                               perl = TRUE)

# SPOTlight
SPOTlight_CoGAPS_deconv <- readRDS(
  file = here::here(sprintf("%s/%s/SPOTlight-CoGAPS_deconv.RDS",
                                  an_tools, robj_dir)))

```

Note on CIBERSORTx we need to run the signature matrix in the CIBERSORTx website since it doesn't work another way.

## Visualization comparison
### Performance comparison
```{r}
tech_ls <- list(
  "MuSiC weighted" = music_mtrx_w,
  "MuSiC all genes" = music_mtrx_a,
  "SPOTlight" = spotlight_mtrx,
  "SCDC" = scdc_deconv,
  "NMFreg" = nmfreg_mtrx,
  "DeconRNAseq" = deconrnaseq_mtrx,
  "RCTD" = rctd_deconv,
  # "deepSCore" = deepSCore_deconv,
  "SPOTlight-CoGAPS" = SPOTlight_CoGAPS_deconv,
  "CIBERSORTx" = CIBERSORTx_deconv)
```

```{r}
source(here::here(glue::glue("{an_tools}/test_synthetic_performance_mod.R")))

output_ls <- lapply(names(tech_ls), function(i) {
  print(i)
  mtrx <- tech_ls[[i]]
  
  if (!is.null(mtrx)) {
    tmp <- test_synthetic_performance(
      test_spots_metadata_mtrx = as.matrix(mtrx[, colnames(synthetic_comp)]),
      spot_composition_mtrx = synthetic_comp)
  } else { tmp <- NULL}
  
  return(tmp)
})

names(output_ls) <- names(tech_ls)
```

### Plot classification metrics
Process the dataframe to get it in tidyverse format
```{r}
metrics_df <- lapply(names(output_ls), function(i) {
  print(i)
  if (!is.null(output_ls[[i]])) {
    tmp_df <- data.frame(output_ls[[i]][1:4])
    tmp_df$Tool <- i
  } else {
    tmp_df <- data.frame(TP = NA, TN = NA, FP = NA, FN = NA)
    tmp_df$Tool <- i
  }
  return(tmp_df)
}) %>% 
  dplyr::bind_rows() %>% 
  dplyr::mutate(
      accuracy = round((TP + TN) / (TP + TN + FP + FN), 2),
      sensitivity = round(TP / (TP + FN), 2),
      specificity = round(TN / (TN + FP), 2),
      precision = round(TP / (TP + FP), 2),
      F1 = round(2 * ((precision * sensitivity) / (precision + sensitivity)), 2)
  ) %>%
  tidyr::pivot_longer(cols = c(accuracy, sensitivity, specificity, F1),
                      names_to = "metric",
                      values_to = "value") %>% 
dplyr::mutate(
  metric = dplyr::if_else(metric == "accuracy", "Accuracy", metric),
  metric = dplyr::if_else(metric == "sensitivity", "Sensitivity", metric),
  metric = dplyr::if_else(metric == "specificity", "Specificity", metric),
  metric = dplyr::if_else(metric == "F1", "F1 score", metric),
)
```

Plot the results
```{r fig.height=6, fig.width=9}
metric_plt <- metrics_df %>%
  dplyr::mutate(
    Tool = factor(x = Tool,
                  levels = techs)
    ) %>%
  ggplot() +
    geom_point(aes(x = Tool,
                   y = value,
                   color = Tool),
               size = 5,
               alpha = 0.9) +
    facet_wrap(. ~ metric) +
    labs(title = "Deconvolution tool benchmarking",
         x = "Tool",
         y = "Metric Value") +
    theme_classic() +
    scale_color_brewer(palette = "Set1") +
    theme(
      strip.text = element_text(hjust = 0.5, size = 14, face = "bold"),
      axis.text.x = element_text(size = 12, angle = 30, hjust = 1),
      axis.text.y = element_text(size = 12),
      axis.title = element_text(size = 15)
      )

metric_plt

cowplot::save_plot(
  filename = here::here(sprintf("%s/%s/class_benchmark.pdf",
                     an_tools, plt_dir)),
  plot = metric_plt,
  base_height = 6,
  base_width = 9)
```

## Compute JSD
Next we're going to compute JSD index for all the predictions to assess how good the predicted proportions are
```{r}
get_jsd <- function(mtrx1, mtrx2) {
  
  ##### Get TRUE JSD between real-predicted proportions #####
  # Initialize matrix
  true_jsd_mtrx <- matrix(nrow = nrow(mtrx1), ncol = 2)
  
  # Loop over all the rows
  for (i in seq_len(nrow(mtrx1))) {

    # Create matrix to feed to JSD
    x <- rbind(mtrx1[i, ],
               mtrx2[i, ])

    # Calculate JSD and save it in true_JSD_mtrx
    if (sum(mtrx2[i, ]) > 0) {
      true_jsd_mtrx[i, 1] <- suppressMessages(JSD(x = x, unit = "log2",
                                                  est.prob = "empirical"))
    } else {
      true_jsd_mtrx[i, 1] <- 1
    }
  }
  
  return(true_jsd_mtrx)
}

JSD_df <- lapply(names(tech_ls), function(i) {
    print(i)
  
    if(!is.null(tech_ls[[i]])) {
      tmp <- get_jsd(mtrx1 = synthetic_comp,
                 mtrx2 = as.matrix(tech_ls[[i]][, colnames(synthetic_comp)]))
      tmp[, 2] <- i
    
    } else {
      tmp <- matrix(data = NA, nrow = 1000, ncol = 2)
      tmp[, 2] <- i
    }
  
  return(data.frame(tmp, stringsAsFactors = FALSE))
}) %>% dplyr::bind_rows()

colnames(JSD_df) <- c("JSD", "Tool")
```

Plot results
```{r fig.height=7, fig.width=11}
JSD_benchmark <- JSD_df %>%
  dplyr::mutate(
    JSD = as.numeric(JSD),
    Tool = factor(x = Tool,
                  levels = techs)) %>%
  ggplot(aes(x = Tool,
             y = JSD,
             fill = Tool,
             color = Tool)) +
  geom_jitter(alpha = 0.5) +
  geom_violin(alpha = 0.8) +
  theme_classic() +
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  theme(
    strip.text = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text.x = element_text(size = 12, angle = 30, hjust = 1),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 15)
  )

JSD_benchmark

cowplot::save_plot(
  filename = here::here(sprintf("%s/%s/JSD_benchmark.pdf",
                     an_tools, plt_dir)),
  plot = JSD_benchmark,
  base_height = 7,
  base_width = 11)
```

Look at dispersion statistics
```{r}
JSD_df %>%
  filter(!is.na(JSD)) %>%
  dplyr::mutate(JSD = as.numeric(JSD)) %>%
  dplyr::group_by(Tool) %>%
  dplyr::summarise(
    Median = round(quantile(JSD, 0.5), 4),
    Q25 = round(quantile(JSD, 0.25), 4),
    Q75 = round(quantile(JSD, 0.75), 4)
  ) %>%
  dplyr::mutate(
    JSD_quantile = sprintf("%s [%s-%s]",
                           Median, Q25, Q75)
  )
```
